---
title: "Syncom community analysis"
author: "Niklas Schandry"
date: "2020"
output: 
  html_document:
    fig_caption: yes
## Widescreen figures
    fig_height: 9
    fig_width: 16
## Highlighting    
    highlight: zenburn
    theme: cosmo
### TOC    
    toc: yes
    toc_float: yes
---

# About

This file documents the community analysis and figure generation for Schandry, et al., 2021, bioRxiv.
Currently this document is subject to change.

# Prepare session

## Libraries and functions

```{r setup, message = F}
library(tidyverse)
library(phyloseq)
library(patchwork)
library(magrittr)
library(ggnetwork)
library(emmeans)
library(vegan)
library(correlation)

source("functions/get_counts_and_overlay.R")
source("functions/phyloseq_corrs.R")
source("functions/overlay_corrnet.R")
source("functions/fix_node_placement.R")
source("functions/compare_conditions.R")
source("functions/read_fasta.R")
```

```{r}
syncom_colors <- c("Random" = "#666699",
                  "Tolerant" = "#9BCB40",
                  "Mixed" = "#B6539F",
                  "Sensitive" = "#FEE377")
```


# Data removal

Replicate 1, appears to have issues that are more pronounced in some samples than in others. This replicate is removed from all Syncoms, leaving four replicates.

# Prepare Data

## General data

```{r}
tax_info <- read_rds("files_publication/taxonomy.rds")
### read growth data
AUC_emmeans <- read_rds("files_publication/AUC_emmeans_pub.rds") %>% 
  left_join(tax_info %>%
            dplyr::select(Genus, Class, Order), by = "Genus")
AUC_vals <- AUC_emmeans %>% 
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(Strain, Chem_conc, AUC) %>%
             pivot_wider(id_cols = Strain, names_from = Chem_conc, values_from = AUC,
                         values_fn = mean) %>%
             as.data.frame() %>% 
             dplyr::mutate(Control = `0_Control0`) %>% 
             dplyr::select("Strain","APO50", "BOA100", "Control" ) 
syncoms_memberships <- read_rds("files_publication/syncom_membership.rds")
```

## Syncom Members

### Tab S1

```{r}
syncoms_memberships %>% 
  group_by(Syncom) %>% 
  arrange(Genus) %>% 
  gt::gt()
```


## Random

```{r read objects for phyloseq random}
# Read sequence table
syncomRandom_noChim    <-  readRDS("files_publication/syncom_full_random_seqtable_nochim.rds")
syncomRandom_tax       <-  dada2::assignTaxonomy(seqs = syncomRandom_noChim,
                                    "files_publication/syncom_full_random_16SrRNA.fasta",
                                     taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomRandom_metadata  <-  readRDS("files_publication/Metadata_publication.rds") %>% 
                              filter(Syncom == "Full_Random")

syncomRandom_strains <-  read_rds("files_publication/syncom_full_random.rds") %>%
  mutate(SynCom = "Random")
```

```{r}
phyl_random <- phyloseq(otu_table(syncomRandom_noChim, taxa_are_rows =F),
                  sample_data(syncomRandom_metadata),
                  tax_table(syncomRandom_tax),
                  phy_tree(read_rds("files_publication/syncom_full_random_GTR_tree.rds")$tree))

phyl_random_glom_strain <- phyl_random %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_random_rel_strain <- transform_sample_counts(phyl_random_glom_strain, function(x) x / sum(x) ) 
```

## Tolerant

```{r read objects for phyloseq tolerant}
# Read sequence table
syncomTolerant_noChim   <-  readRDS("files_publication/syncom_reduced_tolerant_seqtable_nochim.rds")

syncomTolerant_tax      <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_tolerant_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomTolerant_metadata <-  readRDS("files_publication/Metadata_publication.rds") %>%
                          filter(Syncom == "Reduced_Tolerant")

syncomTolerant_strains <-  read_rds("files_publication/syncom_reduced_tolerant.rds") %>%
  mutate(SynCom = "Tolerant")
```

```{r}
phyl_tolerant <- phyloseq(otu_table(syncomTolerant_noChim, taxa_are_rows =F),
                  sample_data(syncomTolerant_metadata),
                  tax_table(syncomTolerant_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_tolerant_GTR_tree.rds")$tree))

phyl_tolerant_glom_strain <- phyl_tolerant %>% 
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_tolerant_rel_strain <- transform_sample_counts(phyl_tolerant_glom_strain, function(x) x / sum(x) ) 
```

## Mixed


```{r read objects for phyloseq mixed}
# Read sequence table
syncomMixed_noChim <-  readRDS("files_publication/syncom_reduced_mixed_seqtable_nochim.rds")
syncomMixed_tax <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_mixed_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomMixed_metadata        <-  readRDS("files_publication/Metadata_publication.rds") %>%                                   filter(Syncom == "Reduced_Mixed")

syncomMixed_strains <-  read_rds("files_publication/syncom_reduced_mixed.rds") %>%
  mutate(SynCom = "Mixed")
```

```{r}
phyl_mixed <- phyloseq(otu_table(syncomMixed_noChim, taxa_are_rows =F),
                  sample_data(syncomMixed_metadata),
                  tax_table(syncomMixed_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_mixed_GTR_tree.rds")$tree))

phyl_mixed_glom_strain <- phyl_mixed %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_mixed_rel_strain <- transform_sample_counts(phyl_mixed_glom_strain, function(x) x / sum(x) ) 
```

## Sensitive


```{r read objects for phyloseq sensitive}
# Read sequence table
syncomSensitive_noChim <-  readRDS("files_publication/syncom_reduced_sensitive_seqtable_nochim.rds")
syncomSensitive_tax <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_sensitive_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomSensitive_metadata        <-  readRDS("files_publication/Metadata_publication.rds") %>%                                   filter(Syncom == "Reduced_Sensitive")

syncomSensitive_strains <-  read_rds("files_publication/syncom_reduced_sensitive.rds") %>%
  mutate(SynCom = "Tolerant")
```

```{r}
phyl_sensitive <- phyloseq(otu_table(syncomSensitive_noChim, taxa_are_rows =F),
                  sample_data(syncomSensitive_metadata),
                  tax_table(syncomSensitive_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_sensitive_GTR_tree.rds")$tree))

phyl_sensitive_glom_strain <- phyl_sensitive %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_sensitive_rel_strain <- transform_sample_counts(phyl_sensitive_glom_strain, function(x) x / sum(x) ) 
```


## Estimate log2Fold changes

```{r}
all_lf2cs_no_time <- bind_rows(list(
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Mixed", Treatment = "APO"),
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Mixed", Treatment = "BOA"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO", 
                     treat2 = "Control", 
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Sensitive", Treatment = "APO"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Sensitive", Treatment = "BOA"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Random", Treatment = "APO"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>% 
  mutate(Syncom = "Random", Treatment = "BOA"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Tolerant", Treatment = "APO"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Tolerant", Treatment = "BOA")))
```

```{r}
all_lf2cs_per_time <- bind_rows(list(
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>% 
  mutate(Syncom = "Mixed", Treatment = "APO"),
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>%
  mutate(Syncom = "Mixed", Treatment = "BOA"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control", 
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>%
  mutate(Syncom = "Sensitive", Treatment = "APO"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment", 
                     treat1 = "BOA", 
                     treat2 = "Control", 
                     alpha = 0.05, 
                     OTUcol = "Strain", 
                     plot = F,
                     return_unfiltered = T,
                     split_var = "Timepoint") %>% 
  mutate(Syncom = "Sensitive", Treatment = "BOA"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO", 
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>%
  mutate(Syncom = "Random", Treatment = "APO"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>% 
  mutate(Syncom = "Random", Treatment = "BOA"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T,
                     split_var = "Timepoint") %>% 
  mutate(Syncom = "Tolerant", Treatment = "APO"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA", 
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T,
                     split_var = "Timepoint") %>%
  mutate(Syncom = "Tolerant", Treatment = "BOA")))
```

# Beta Diverstiy


### CCA with jaccard

```{r}
cca_jaccard_random_rel_strain <-  phyl_random_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_random_plot<- plot_ordination(phyl_random_rel_strain,
                cca_jaccard_random_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Random Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_tolerant_rel_strain <-  phyl_tolerant_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_tolerant_plot <- plot_ordination(phyl_tolerant_rel_strain,
                cca_jaccard_tolerant_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Tolerant Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_mixed_rel_strain <-  phyl_mixed_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_mixed_plot <- plot_ordination(phyl_mixed_rel_strain,
                cca_jaccard_mixed_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Mixed Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_sensitive_rel_strain <-  phyl_sensitive_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_sensitve_plot <- plot_ordination(phyl_sensitive_rel_strain,
                cca_jaccard_sensitive_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Sensitive Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  
```

#### Fig2

```{r}
(cca_jaccard_random_plot | cca_jaccard_tolerant_plot) /
(cca_jaccard_mixed_plot | cca_jaccard_sensitve_plot) +
  plot_layout(guides = "collect",
              heights = c(1,1)) +
  plot_annotation(tag_levels = "A", caption = "Formula: Treatment + Timepoint") &
  theme(text=element_text(family = "NimbusMon"), strip.background = element_rect(fill = "white"))
```


## PERMANOVA

```{r}
random_bray <- phyl_random_rel_strain %>% phyloseq::distance("bray")
random_jaccard <- phyl_random_rel_strain %>% phyloseq::distance("jaccard")
random_wuni <- phyl_random_rel_strain %>% phyloseq::distance("wunifrac")
random_samples <- phyl_random_rel_strain %>% sample_data() %>% as("data.frame")
```

```{r}
vegan::adonis(random_bray ~ Timepoint + Treatment,data = random_samples )
```
```{r}
vegan::adonis(random_wuni ~ Timepoint + Treatment,data = random_samples )
```
```{r}
vegan::adonis(random_jaccard ~ Timepoint + Treatment,data = random_samples )
```
```{r}
tolerant_bray <- phyl_tolerant_rel_strain %>% phyloseq::distance("bray")
tolerant_jaccard <- phyl_tolerant_rel_strain %>% phyloseq::distance("jaccard")
tolerant_wuni <- phyl_tolerant_rel_strain %>% phyloseq::distance("wunifrac")
tolerant_samples <- phyl_tolerant_rel_strain %>% sample_data() %>% as("data.frame")
```

```{r}
vegan::adonis(tolerant_bray ~ Timepoint + Treatment,data = tolerant_samples )
```
```{r}
vegan::adonis(tolerant_jaccard ~ Timepoint + Treatment,data = tolerant_samples )
```
```{r}
vegan::adonis(tolerant_wuni ~ Timepoint + Treatment,data = tolerant_samples )
```
```{r}
mixed_bray <- phyl_mixed_rel_strain %>% phyloseq::distance("bray")
mixed_jaccard <- phyl_mixed_rel_strain %>% phyloseq::distance("jaccard")
mixed_wuni <- phyl_mixed_rel_strain %>% phyloseq::distance("wunifrac")
mixed_samples <- phyl_mixed_rel_strain %>% sample_data() %>% as("data.frame")
```

```{r}
vegan::adonis(mixed_bray ~ Timepoint + Treatment,data = mixed_samples )
```
```{r}
vegan::adonis(mixed_jaccard ~ Timepoint + Treatment,data = mixed_samples )
```
```{r}
vegan::adonis(mixed_wuni ~ Timepoint + Treatment,data = mixed_samples )
```
```{r}
sensitive_bray <- phyl_sensitive_rel_strain %>% phyloseq::distance("bray")
sensitive_jaccard <- phyl_sensitive_rel_strain %>% phyloseq::distance("jaccard")
sensitive_wuni <- phyl_sensitive_rel_strain %>% phyloseq::distance("wunifrac")
sensitive_samples <- phyl_sensitive_rel_strain %>% sample_data() %>% as("data.frame")
```

```{r}
vegan::adonis(sensitive_bray ~ Timepoint + Treatment,data = sensitive_samples )
```
```{r}
vegan::adonis(sensitive_jaccard ~ Timepoint + Treatment,data = sensitive_samples )
```
```{r}
vegan::adonis(sensitive_wuni ~ Timepoint + Treatment,data = sensitive_samples )
```

### Tab S1 PERMANOVA

```{r}
bind_rows(list(vegan::adonis(random_jaccard ~ Timepoint * Treatment, data = random_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Random"),
vegan::adonis(tolerant_jaccard ~ Timepoint * Treatment, data = tolerant_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Tolerant"),
vegan::adonis(mixed_jaccard ~ Timepoint * Treatment, data = mixed_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Mixed"),
vegan::adonis(sensitive_jaccard ~ Timepoint * Treatment, data = sensitive_samples ) %>% 
  .$aov.tab %>% 
  as.tibble(rownames= "term") %>% 
  mutate(Syncom = "Sensitive"))) %>%
  {set_colnames(.,c(colnames(.)[1:6], "p.value", "Syncom"))} %>% 
  mutate(round(across(where(is.numeric)),4)) %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S5. PERMANOVA", subtitle = "999 Permutations") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","R2","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  gt::gtsave("TabS5_permanova.html")
```

```{r}
bind_rows(list(
  vegan::adonis(random_jaccard ~ Timepoint * Treatment, data = random_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Random"),
vegan::adonis(tolerant_jaccard ~ Timepoint * Treatment, data = tolerant_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Tolerant"),
vegan::adonis(mixed_jaccard ~ Timepoint * Treatment, data = mixed_samples ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Mixed"),
vegan::adonis(sensitive_jaccard ~ Timepoint * Treatment, data = sensitive_samples ) %>% 
  .$aov.tab %>% 
  as.tibble(rownames= "term") %>% 
  mutate(Syncom = "Sensitive"))) %>%
  {set_colnames(.,c(colnames(.)[1:6], "p.value", "Syncom"))} %>% 
  mutate(round(across(where(is.numeric)),4)) %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
#  gt::tab_header("Table S5. PERMANOVA", subtitle = "999 Permutations") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","R2","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("TabS5_permanova_gt.rds")
```

### PERMANOVA Subsets

```{r}
random_jaccard_APO_DMSO <- phyl_random_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  phyloseq::distance("jaccard")

random_jaccard_BOA_DMSO <- phyl_random_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
  phyloseq::distance("jaccard")

tolerant_jaccard_APO_DMSO <- phyl_tolerant_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  phyloseq::distance("jaccard")

tolerant_jaccard_BOA_DMSO <- phyl_tolerant_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
  phyloseq::distance("jaccard")

mixed_jaccard_APO_DMSO <- phyl_mixed_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  phyloseq::distance("jaccard")

mixed_jaccard_BOA_DMSO <- phyl_mixed_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
  phyloseq::distance("jaccard")

sensitive_jaccard_APO_DMSO <- phyl_sensitive_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  phyloseq::distance("jaccard")

sensitive_jaccard_BOA_DMSO <- phyl_sensitive_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
  phyloseq::distance("jaccard")
```

```{r}
random_jaccard_APO_DMSO_samp <- phyl_random_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  sample_data() %>% as("data.frame")

random_jaccard_BOA_DMSO_samp <- phyl_random_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
 sample_data() %>% as("data.frame")

tolerant_jaccard_APO_DMSO_samp <- phyl_tolerant_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  sample_data() %>% as("data.frame")

tolerant_jaccard_BOA_DMSO_samp <- phyl_tolerant_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
 sample_data() %>% as("data.frame")

mixed_jaccard_APO_DMSO_samp <- phyl_mixed_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  sample_data() %>% as("data.frame")

mixed_jaccard_BOA_DMSO_samp <- phyl_mixed_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
 sample_data() %>% as("data.frame")

sensitive_jaccard_APO_DMSO_samp <- phyl_sensitive_rel_strain %>%
  subset_samples(Treatment != "BOA") %>% 
  sample_data() %>% as("data.frame")

sensitive_jaccard_BOA_DMSO_samp <- phyl_sensitive_rel_strain %>%
  subset_samples(Treatment != "APO") %>% 
 sample_data() %>% as("data.frame")
```

```{r}
treat_permaov_tab <- bind_rows(list(
  vegan::adonis(random_jaccard_APO_DMSO ~ Treatment, data = random_jaccard_APO_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Random",
         Treatment = "APO")  ,
vegan::adonis(random_jaccard_BOA_DMSO ~ Treatment, data = random_jaccard_BOA_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Random",
         Treatment = "BOA"),
  vegan::adonis(tolerant_jaccard_APO_DMSO ~ Treatment, data = tolerant_jaccard_APO_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Tolerant",
         Treatment = "APO")  ,
vegan::adonis(tolerant_jaccard_BOA_DMSO ~ Treatment, data = tolerant_jaccard_BOA_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Tolerant",
         Treatment = "BOA"),
  vegan::adonis(mixed_jaccard_APO_DMSO ~ Treatment, data = mixed_jaccard_APO_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Mixed",
         Treatment = "APO")  ,
vegan::adonis(mixed_jaccard_BOA_DMSO ~ Treatment, data = mixed_jaccard_BOA_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Mixed",
         Treatment = "BOA"),
  vegan::adonis(sensitive_jaccard_APO_DMSO ~ Treatment, data = sensitive_jaccard_APO_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Sensitive",
         Treatment = "APO")  ,
vegan::adonis(sensitive_jaccard_BOA_DMSO ~ Treatment, data = sensitive_jaccard_BOA_DMSO_samp ) %>% 
  .$aov.tab %>% 
  as_tibble(rownames= "term") %>% 
  mutate(Syncom = "Sensitive",
         Treatment = "BOA"))) %>%
  {set_colnames(.,c(colnames(.)[1:6], "p.value", "Syncom", "Treatment"))} %>% 
  mutate(round(across(where(is.numeric)),4))
```

### Table

```{r}
treat_permaov_tab %>% 
  group_by(Syncom, Treatment) %>% 
  gt::gt() %>% 
#  gt::tab_header("Table S5. PERMANOVA", subtitle = "999 Permutations") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","R2","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("files_publication/TabS3_PERMANOVA_per_treatment_gt.rds")
```


# Alpha Diversity 

## Observed

### Plots

```{r}
observed_random <- phyl_random_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Observed") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_"))

observed_random_plot <- observed_random %>% 
  ggplot(aes(x = Treatment, y = Observed, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,15.5)) +
  ggtitle(paste("Random Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

observed_tolerant <- phyl_tolerant_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Observed") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>%
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_")) 

observed_tolerant_plot <- observed_tolerant %>% 
  ggplot(aes(x = Treatment, y = Observed, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,15.5)) +
  ggtitle(paste("Tolerant Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

observed_mixed <- phyl_mixed_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Observed") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_")) 

observed_mixed_plot <- observed_mixed %>% 
  ggplot(aes(x = Treatment, y = Observed, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,15.5)) +
  ggtitle(paste("Mixed Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

observed_sensitive <- phyl_sensitive_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Observed") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_")) 

observed_sensitive_plot <- observed_sensitive %>% 
  ggplot(aes(x = Treatment, y = Observed, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,15.5)) +
  ggtitle(paste("Sensitive Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

#### Fig S2


```{r}
FigS2 <- ((observed_random_plot | observed_tolerant_plot ) / 
   (observed_mixed_plot |observed_sensitive_plot ) ) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon")) 
```

```{r}
FigS2
```
```{r}
png("Supplemental_Figures/Figure_S2.png",width = 720, height = 560)
FigS2
dev.off()
```


### AOV

```{r}
observed_aov <- bind_rows(list(
  observed_random %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Random"),
  observed_tolerant %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Tolerant"),
  observed_mixed %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Mixed"),
  observed_sensitive %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Sensitive")))
```

```{r}
observed_aov %>% write_rds("rds/observed_aov.rds")
```

#### Tab S3


```{r}
observed_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S3. ANOVA: Observed Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) 
```
```{r}
observed_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S3. ANOVA: Observed Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  gt::gtsave("TabS1_observed_aov.html")
```
```{r}
observed_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("TabS1_gt.rds")
```

### Pairwise CLD

```{r}
observed_all_cld <- bind_rows(list(
observed_random %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Random"),
observed_tolerant %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Tolerant"),
observed_mixed %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Mixed"),
observed_sensitive %>% 
  {lm(Observed ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Sensitive"))) %>% 
  mutate(Group = str_extract(Group, "[a-z]+")) %>% 
  mutate("95% CI" = paste0("[", 
                           round(as.numeric(Conf.lower), 3),
                           " - ",
                           round(as.numeric(Conf.upper), 3),
                           "]")) %>% 
  dplyr::select(Syncom, Timepoint, Treatment, estimate, `95% CI`, Group)
```

#### Tab S4

```{r}
observed_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
  gt::tab_header("Table S4. Observed diversity means and pairwise comparisons") %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels("Group"))
```
```{r}
observed_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
  gt::tab_header("Table S4. Observed diversity means and pairwise comparisons") %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels("Group")) %>% 
  gt::gtsave("TabS4_observed_pairwise_cld.html")
```

```{r}
observed_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels("Group")) %>% 
  write_rds("TabS2_gt.rds")
```

## Shannon

### Plots

```{r}
## Random

shannon_random <- phyl_random_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Shannon") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_"))

shannon_random_plot <- shannon_random %>% 
  ggplot(aes(x = Treatment, y = Shannon, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,1.55)) +
  ggtitle(paste("Random Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

## Tolerant

shannon_tolerant <- phyl_tolerant_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Shannon") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_")) 


shannon_tolerant_plot <- shannon_tolerant %>% 
  ggplot(aes(x = Treatment, y = Shannon, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,1.55)) +
  ggtitle(paste("Tolerant Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

## Mixed

shannon_mixed <- phyl_mixed_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Shannon") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_"))

shannon_mixed_plot <- shannon_mixed %>% 
  ggplot(aes(x = Treatment, y = Shannon, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,1.55)) +
  ggtitle(paste("Mixed Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

## Sensitive

shannon_sensitive <- phyl_sensitive_glom_strain %>% 
  phyloseq::estimate_richness(measures = "Shannon") %>% 
  rownames_to_column("sample") %>% 
  separate(sample,
           c("Type", "Assembly","Treatment", NA, NA, "Rep", "Timepoint"),
           sep = "_") %>% 
  mutate(Treatment = case_when(Treatment == "DMSO" ~ "Control",
                               TRUE ~ Treatment)) %>% 
  mutate(Syncom = paste(Type, Assembly, sep = "_"))

shannon_sensitive_plot <- shannon_sensitive %>% 
  ggplot(aes(x = Treatment, y = Shannon, color = Treatment)) +
  geom_boxplot() +
  facet_grid(~Timepoint) +
  geom_point() +
  lims(y = c(0,1.55)) +
  ggtitle(paste("Sensitive Syncom")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```

Phyloseq raises a warning regarding lack of singletons. Shannon diversity does rely on low-abundance species, since these communities are in-vitro, singletons are not really expected.

#### Fig S3

```{r}
FigS3 <- ((shannon_random_plot | shannon_tolerant_plot) /
  ( shannon_mixed_plot | shannon_sensitive_plot ) ) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon")) 
```

```{r}
FigS3
```

```{r}
png("Supplemental_Figures/Figure_S3.png",width = 720, height = 560)
FigS3
dev.off()
```

### AOV

```{r}
bind_rows(list(shannon_random %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Random"),
  shannon_tolerant %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Tolerant"),
  shannon_mixed %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Mixed"),
  shannon_sensitive %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Sensitive"))) %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S5. ANOVA: Shannon Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  )
```
```{r}
bind_rows(list(shannon_random %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Random"),
  shannon_tolerant %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Tolerant"),
  shannon_mixed %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Mixed"),
  shannon_sensitive %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Sensitive"))) %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  #gt::tab_header("Table S3. ANOVA: Shannon Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("TabS3_gt.rds")
```

```{r}
shannon_aov <- bind_rows(list(
  shannon_random %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Random"),
  shannon_tolerant %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Tolerant"),
  shannon_mixed %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Mixed"),
  shannon_sensitive %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  aov %>% 
  broom::tidy() %>% 
  mutate(Syncom = "Sensitive")))
```

```{r}
shannon_aov %>% write_rds("rds/shannon_aov.rds")
```

#### Tab S5

```{r}
shannon_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S5. ANOVA: Shannon Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) 
```
```{r}
shannon_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::tab_header("Table S5. ANOVA: Shannon Diversity") %>%  
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  gt::gtsave("TabS5_shannon_aov.html")
```
```{r}
shannon_aov %>% 
  group_by(Syncom) %>% 
  gt::gt() %>% 
  gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("term","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("TabS3_gt.rds")
```

### Pairwise CLD

```{r}
shannon_all_cld <- bind_rows(list(
shannon_random %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Random"),
shannon_tolerant %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Tolerant"),
shannon_mixed %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Mixed"),
shannon_sensitive %>% 
  {lm(Shannon ~ Treatment*Timepoint, data = .)} %>% 
  emmeans::emmeans( ~ Treatment, by = "Timepoint") %>% 
  emmeans::CLD(Letters = letters) %>%
  broom::tidy() %>% 
  {set_colnames(., c(colnames(.)[1:2],
                    "estimate",
                    "SE",
                    "Conf.lower",
                    "Conf.upper",
                    "Group"))} %>% 
  # Drop SE column since the model SE is not so interesting
    dplyr::select(-SE) %>% 
  mutate(Syncom = "Sensitive"))) %>% 
  mutate(Group = str_extract(Group, "[a-z]+")) %>% 
  mutate("95% CI" = paste0("[", 
                           round(as.numeric(Conf.lower), 3),
                           " - ",
                           round(as.numeric(Conf.upper), 3),
                           "]")) %>% 
  dplyr::select(Syncom, Timepoint, Treatment, estimate, `95% CI`, Group)
```


```{r}
shannon_all_cld %>% write_rds("rds/shannon_all_cld.rds")
```

#### Tab S4

```{r}
shannon_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
  gt::tab_header("Table S4. Shannon diversity means and pairwise comparisons") %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels(
      "Group"))
```
```{r}
shannon_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
  gt::tab_header("Table S4. Shannon diversity means and pairwise comparisons") %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels(
      "Group")) %>% 
  gt::gtsave("TabS4_shannon_pairwise.html")
```
```{r}
shannon_all_cld %>% 
  mutate(estimate = round(estimate, 3)) %>% 
  group_by(Syncom, Timepoint) %>% 
  gt::gt()  %>% 
#  gt::tab_header("Table S4. Shannon diversity means and pairwise comparisons") %>% 
  gt::tab_style(style = gt::cell_fill(color = "#cbe0ff"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "a" )) %>% 
  gt::tab_style(style = gt::cell_fill(color = "#ddc3d0"),
      locations = gt::cells_body(
      columns = c("Group"),
      rows = Group == "b")) %>%   
  gt::tab_footnote(
    footnote = "Different letters indicate significant differences within one Syncom / Timepoint combination (alpha = 0.05), Tukey adjusted",
    locations = gt::cells_column_labels(
      "Group")) %>% 
  write_rds("TabS4_gt.rds")
```

# AUC vs relative Abundance

```{r}
strain_AUC <- AUC_emmeans %>%
  filter(Treat.x %in% c("Control","50uM_APO","100uM_BOA")) %>%
  mutate(Treatment = case_when(Treat.x == "Control" ~ Treat.x,
                               Treat.x == "50uM_APO" ~ "APO",
                               Treat.x == "100uM_BOA" ~ "BOA")) %>% 
  dplyr::select(Strain, Genus ,Treatment, AUC, estimate) %>%
  group_by(Strain,Treatment) %>%
  summarize(mean_AUC = mean(AUC, na.rm = T),
            mean_rel_AUC = mean(estimate, na.rm = T)) %>%
  ungroup  
```

## Plot l2FC community vs AUC ratio

```{r}
growth_data_l2fc <- left_join(strain_AUC, all_lf2cs_no_time,  by = c("Strain", "Treatment")) %>% 
  filter(!is.na(Treatment),
         !is.na(Syncom),
         !is.na(log2FoldChange),
         Treatment != "Control") 
```

```{r}
relabund_vs_auc <- rbind(
 phyl_random_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Strain, Treatment, Abundance, Syncom) %>% 
  group_by(Strain, Treatment, Syncom) %>% 
  summarize(mean_abundance = mean(Abundance)) %>% 
   ungroup(),
 phyl_tolerant_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Strain, Treatment, Abundance, Syncom) %>% 
  group_by(Strain, Treatment, Syncom) %>% 
  summarize(mean_abundance = mean(Abundance)) %>% 
   ungroup(),
 phyl_sensitive_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Strain, Treatment, Abundance, Syncom) %>% 
  group_by(Strain, Treatment, Syncom) %>% 
  summarize(mean_abundance = mean(Abundance)) %>% 
   ungroup(),
 phyl_mixed_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Strain, Treatment, Abundance, Syncom) %>% 
  group_by(Strain, Treatment, Syncom) %>% 
  summarize(mean_abundance = mean(Abundance)) %>% 
   ungroup()) %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment"))
 
```
```{r}
lm(mean_abundance ~ mean_AUC:Treatment, data = relabund_vs_auc) %>% summary
```


## Estimate R^2

### BOA

```{r}
lm(log2FoldChange ~ mean_rel_AUC + Syncom, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "BOA")) %>% 
  summary()
```

```{r}
r2_boa <-  lm(log2FoldChange ~ mean_rel_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "BOA")) %>% 
  summary() %>% .$adj.r.squared
```

No correlation. 

```{r}
lm(log2FoldChange ~ mean_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "BOA")) %>% 
  summary()
```

```{r}
r2_boa_absAUC <-  lm(log2FoldChange ~ mean_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "BOA")) %>% 
  summary() %>%
  .$adj.r.squared
```

### APO

```{r}
lm(log2FoldChange ~ mean_rel_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "APO")) %>% 
  summary()
r2_apo <-  lm(log2FoldChange ~ mean_rel_AUC, 
              data = growth_data_l2fc %>% 
              filter(Treatment == "APO")) %>% 
  summary() %>%
  .$adj.r.squared
```


```{r}
lm(log2FoldChange ~ mean_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "APO")) %>% 
  summary()
r2_apo_absAUC <- lm(log2FoldChange ~ mean_AUC, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "APO")) %>% 
  summary() %>% .$adj.r.squared
```

No correlation. 

```{r}
lm(log2FoldChange ~ mean_rel_AUC + Syncom, 
   data = growth_data_l2fc %>% 
     filter(Treatment == "APO")) %>% 
  summary()
```

#### Tab S8

```{r}
r2adj <-tibble(r2 = c(r2_apo, r2_boa, r2_apo_absAUC, r2_boa_absAUC) ,Treatment = c("APO","BOA","APO","BOA"), AUC = c("relative","relative","absolute","absolute"))
r2adj %>% write_rds("files_publication/rsquare_l2FC_AUC.rds")
```

```{r}
r2adj %>% gt::gt()
```

## Plot AUC vs l2fc

```{r}
plot_dotplot_auc_vs_l2fc <- growth_data_l2fc %>% 
  ggplot(aes(x = mean_rel_AUC,
             y = log2FoldChange)) +
  geom_vline(aes(xintercept = 1), color = "darkgrey") +
  geom_hline(aes(yintercept = 0), color = "darkgrey") +
  geom_smooth(method = "lm", color = "black", se = T) +
  geom_point(colour = "black", 
             size = 5, 
             alpha = 0.7) +
  geom_point(aes(color = Syncom),
             size = 5, 
             alpha = 0.7) +
  scale_color_manual(values = syncom_colors) +
  facet_wrap(~Treatment, scales = "free_x") +
  theme_bw(base_family = "NimbusMon") +
  xlab(paste("Effect on strain grown in isolation \n [Ratio area under the curve (Treatment / Control)]")) +
  ylab(paste("Effect on strain grown in community \n [log2Foldchange (Treatment / Control)]")) +
  ggtitle("Growth and abundance changes by treatment") +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        text = element_text(family = "NimbusMon"),
        strip.background = element_blank(),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",
                                  face = "bold"),
        legend.position = "none") 
plot_dotplot_auc_vs_l2fc
```
```{r}
plot_dotplot_abs_auc_vs_l2fc <- growth_data_l2fc %>% 
  ggplot(aes(x = mean_AUC,
             y = log2FoldChange)) +
 # geom_vline(aes(xintercept = 1), color = "darkgrey") +
  geom_hline(aes(yintercept = 0), color = "darkgrey") +
  geom_smooth(method = "lm", color = "black", se = T) +
  geom_point(colour = "black", 
             size = 5, 
             alpha = 0.7) +
  geom_point(aes(color = Syncom),
             size = 5, 
             alpha = 0.7) +
  scale_color_manual(values = syncom_colors) +
  facet_wrap(~Treatment, scales = "free_x") +
  theme_bw(base_family = "NimbusMon") +
  xlab(paste("Growth in Treatment")) +
  ylab(paste("Effect on strain grown in community \n [log2Foldchange (Treatment / Control)]")) +
  ggtitle("Growth and abundance changes by treatment") +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        text = element_text(family = "NimbusMon"),
        strip.background = element_blank(),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",
                                  face = "bold"),
        legend.position = "none") 
plot_dotplot_abs_auc_vs_l2fc
```

## Strains across communities

```{r}
all_lf2cs_no_time %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment")) %>% 
  mutate(genus_strain = str_c(Genus,Strain, sep = "\n")) %>% 
  group_by(Strain) %>% 
  add_count(Treatment) %>% 
  filter(n > 1) %>% 
  ggplot(aes( x = mean_rel_AUC,y=log2FoldChange, shape = Treatment)) +
  facet_wrap(~genus_strain, scales = "free_y") +
  geom_point(size = 4) +
  geom_point(aes(color = Syncom), size = 3) +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.background = element_rect(fill = "white")) +
  scale_color_manual(values = syncom_colors) +
  labs(title = "Changes in abundance",
       sub = "Strains included in more than one community",
       y = "log2FC in Community\n(across all timepoints)",
       x = expression("AUC(Treatment) / AUC(Control)"))
  #ggthemes::scale_color_few()
```

```{r}
all_lf2cs_no_time %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment")) %>% 
  mutate(genus_strain = str_c(Genus,Strain, sep = "\n")) %>% 
  group_by(Strain) %>% 
  add_count(Treatment) %>% 
  filter(n > 1) %>% 
  ggplot(aes( x = Treatment,y=log2FoldChange)) +
  facet_wrap(~genus_strain, scales = "free_y") +
  geom_point(size = 4) +
  geom_point(aes(color = Syncom), size = 3) +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.background = element_rect(fill = "white")) +
  scale_color_manual(values = syncom_colors) +
  labs(title = "Changes in abundance",
       sub = "Strains included in more than one community",
       y = "log2FC in Community\n(across all timepoints)",
       x = "Treatment")
  #ggthemes::scale_color_few()
```

```{r}
all_lf2cs_no_time %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment")) %>% 
  mutate(genus_strain = str_c(Genus,Strain, sep = "\n")) %>% 
  group_by(Strain) %>% 
  add_count(Treatment) %>% 
  filter(n > 1) %>% 
  ggplot(aes( y=log2FoldChange, x = Treatment, group = Syncom)) +
  facet_wrap(~genus_strain, ncol = 5, scales = "free_x") +
  geom_hline(aes(yintercept = 0),
             linetype = "dashed",
             color = "darkgrey" ) +
  geom_errorbar(aes(ymin = log2FoldChange-lfcSE, 
                      ymax = log2FoldChange + lfcSE),
                 position = position_dodge(width = 0.6),
                 width = 0.2) +
  geom_point(size = 4,
             position = position_dodge(width = 0.6)) +
  geom_point(aes(color = Syncom),
             size = 3,
             position = position_dodge(width = 0.6)) +
  geom_vline(aes(xintercept = 1.5)) +
  scale_x_discrete(limits = rev) + 
  coord_flip() +
  theme_bw() +
  theme(text = element_text(family = "Arial"),
        strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_color_manual(values = syncom_colors) +
  labs(title = "Changes in abundance",
       sub = "Strains included in more than one community",
       y = "log2FC vs Control (across all timepoints)",
       x = "Treatment")
  #ggthemes::scale_color_few()
```

#### Per timepoint

```{r}
plot_l2fc_per_community <-all_lf2cs_per_time %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment")) %>% 
  mutate(genus_strain = str_c(Genus,Strain, sep = "\n")) %>% 
  group_by(Strain, Timepoint) %>% 
  add_count(Treatment) %>% 
  filter(n > 1) %>% 
  ggplot(aes( y=log2FoldChange, x = interaction(Timepoint, Treatment), group = Syncom)) +
  facet_wrap(~genus_strain, ncol = 3) +
  geom_hline(aes(yintercept = 0),
             linetype = "dashed",
             color = "darkgrey" ) +
  geom_errorbar(aes(ymin = log2FoldChange-lfcSE, 
                      ymax = log2FoldChange + lfcSE),
                 position = position_dodge(width = 0.6),
                 width = 0.2) +
  geom_point(size = 4,
             position = position_dodge(width = 0.6)) +
  geom_point(aes(color = Syncom),
             size = 3,
             position = position_dodge(width = 0.6)) +
  geom_vline(aes(xintercept = 4.5)) +
  scale_x_discrete(limits = rev) + 
  coord_flip() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.background = element_rect(fill = "white")) +
  scale_color_manual(values = syncom_colors) +
  labs(title = "Changes in abundance",
       sub = "Strains included in more than one community",
       y = "log2FC in Community (across all timepoints)",
       x = "Treatment")
  #ggthemes::scale_color_few()
plot_l2fc_per_community
```

### Figure

```{r}
plot_dotplot_auc_vs_l2fc / plot_l2fc_per_community +
  plot_layout(heights = c(3,5),
              guides = "collect") &
  theme(legend.position = "bottom")
  
```



# Relative abundance changes

## Calculate l2FC

### Random

```{r}
## APO /BOA
APO_BOA_random <- bind_rows(
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>%   mutate(Timepoint = "24h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## APO / Control

APO_DMSO_random <- bind_rows(
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## BOA / Control

BOA_DMSO_random <- bind_rows(
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_random_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))
```


### Tolerant

```{r}
## APO /BOA

APO_BOA_tolerant <- bind_rows(
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## APO / Control

APO_DMSO_tolerant <- bind_rows(
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## BOA / Control

BOA_DMSO_tolerant <- bind_rows(
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_tolerant_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))
```


### Mixed

```{r}
## APO /BOA

APO_BOA_mixed <- bind_rows(
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## APO / Control

APO_DMSO_mixed <- bind_rows(
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## BOA / Control

BOA_DMSO_mixed <- bind_rows(
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_mixed_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))
```


### Sensitive

```{r}
## APO /BOA

APO_BOA_sensitive <- bind_rows(
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "BOA", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## APO / Control

APO_DMSO_sensitive <- bind_rows(
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "APO", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))

## BOA / Control

BOA_DMSO_sensitive <- bind_rows(
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "24h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "24h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "48h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "48h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "72h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "72h"),
phyl_sensitive_glom_strain %>%
  subset_samples(Timepoint == "96h") %>% 
  compare_conditions(treatvar = "Treatment", treat1= "BOA", treat2 = "Control", plot = FALSE) %>% 
  mutate(Timepoint = "96h"))
```

## Fig S3

```{r}
colors_l2fc <- wesanderson::wes_palette("Moonrise2", 3, "discrete")
```


```{r}
plot_sens_allcomparisons <- (bind_rows(
  list(APO_DMSO_sensitive %>% mutate(Comparison = "APO / Control"),
       APO_BOA_sensitive %>% mutate(Comparison = "APO / BOA"),
       BOA_DMSO_sensitive %>% mutate(Comparison = "BOA / Control"))) %>% 
  ggplot(aes( x = log2FoldChange, y = fct_rev(Genus), color = Comparison)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), position = "dodge",width = 0.2) +
  geom_point() +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), width = 0.2) +
  geom_vline(aes(xintercept = 0), lty = 3) +
  ylab(element_blank()) +
  facet_wrap(~Timepoint) +
 #ggtitle("Sensitive: log2FC") +
  scale_y_discrete(label=abbreviate) +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon")) +
  scale_color_manual(values = colors_l2fc) +
  labs(x = "log2FoldChange (sensitive)") ) /
  (phyl_sensitive_rel_strain %>% 
  psmelt() %>%
  filter(Genus %in% c(APO_DMSO_sensitive$Genus,APO_BOA_sensitive$Genus, BOA_DMSO_sensitive$Genus)) %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  #ggtitle(paste("Sensitive: Relative Genus abundances")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y"))
```

```{r}
plot_mix_allcomparisons <- (bind_rows(
  list(APO_DMSO_mixed %>% mutate(Comparison = "APO / Control"),
       APO_BOA_mixed %>% mutate(Comparison = "APO / BOA"),
       BOA_DMSO_mixed %>% mutate(Comparison = "BOA / Control"))) %>% 
  ggplot(aes( x = log2FoldChange, y = fct_rev(Genus), color = Comparison)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), position = "dodge",width = 0.2) +
  geom_point() +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), width = 0.2) +
  geom_vline(aes(xintercept = 0), lty = 3) +
  ylab(element_blank()) +
  facet_wrap(~Timepoint) +
  #ggtitle("Mixed: log2FC") +
  scale_y_discrete(label=abbreviate) +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon")) +
  scale_color_manual(values = colors_l2fc) +
  labs(x = "log2FoldChange (mixed)") ) /
  (phyl_mixed_rel_strain %>% 
  psmelt() %>%
  filter(Genus %in% c(APO_DMSO_mixed$Genus,APO_BOA_mixed$Genus, BOA_DMSO_mixed$Genus)) %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() +
  scale_y_continuous(n.breaks = 4) +
  #ggtitle(paste("Mixed: Relative Genus abundances")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y"))
```

```{r}
plot_tol_allcomparisons <- (bind_rows(
  list(APO_DMSO_tolerant %>% mutate(Comparison = "APO / Control"),
       APO_BOA_tolerant %>% mutate(Comparison = "APO / BOA"),
       BOA_DMSO_tolerant %>% mutate(Comparison = "BOA / Control"))) %>% 
  filter(Genus != "Nocardioides") %>%
  ggplot(aes( x = log2FoldChange, y = fct_rev(Genus), color = Comparison)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), position = "dodge",width = 0.2) +
  geom_point() +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), width = 0.2) +
  geom_vline(aes(xintercept = 0), lty = 3) +
  ylab(element_blank()) +
  facet_wrap(~Timepoint) +
  #ggtitle("Tolerant: log2FC") +
  scale_y_discrete(label=abbreviate) +   
  theme_bw() +
  theme(text = element_text(family = "NimbusMon")) +
  scale_color_manual(values = colors_l2fc) +
  labs(x = "log2FoldChange (tolerant)") ) /
  (phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  filter(Genus %in% c(APO_DMSO_tolerant$Genus, APO_BOA_tolerant$Genus, BOA_DMSO_tolerant$Genus),
         Genus != "Nocardioides") %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  #ggtitle(paste("Tolerant: Relative Genus abundances")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y",nrow = 2))
```


```{r}
plot_rand_allcomparisons <- (bind_rows(
  list(APO_DMSO_random %>% mutate(Comparison = "APO / Control"),
       APO_BOA_random %>% mutate(Comparison = "APO / BOA"),
       BOA_DMSO_random %>% mutate(Comparison = "BOA / Control"))) %>% 
  ggplot(aes( x = log2FoldChange, y = fct_rev(Genus), color = Comparison)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), position = "dodge",width = 0.2) +
  geom_point() +
  geom_errorbar(aes(xmin = log2FoldChange - lfcSE, xmax = log2FoldChange + lfcSE), width = 0.2) +
  geom_vline(aes(xintercept = 0), lty = 3) +
  ylab(element_blank()) +
  facet_wrap(~Timepoint) +
  #ggtitle("Random: log2FC") +
  scale_y_discrete(label=abbreviate) +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon")) +
  scale_color_manual(values = colors_l2fc) +
  labs(x = "log2FoldChange (random)") ) /
  (phyl_random_rel_strain %>% 
  psmelt() %>%
  filter(Genus %in% c(APO_DMSO_random$Genus, APO_BOA_random$Genus, BOA_DMSO_random$Genus)) %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  #ggtitle(paste("Random: Relative Genus abundances")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y",nrow = 2))
```

### Fig S3 - Plot

```{r}
((plot_rand_allcomparisons | plot_tol_allcomparisons) / (plot_mix_allcomparisons | plot_sens_allcomparisons)) +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = 'collect' ) &
  theme(text = element_text(family ="NimbusMon"),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 8),
        legend.position = "bottom") 
```

# Overlap Random and Tolerant

```{r}
tol_strains <- phyl_tolerant_rel_strain %>% 
  psmelt() %>% .$Strain
rand_strains <- phyl_random_rel_strain %>% 
  psmelt() %>% .$Strain
rand_tol_genusstrain <- rbind(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  dplyr::select(Strain, Genus_Strain),
  phyl_random_rel_strain %>% 
  psmelt() %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  dplyr::select(Strain, Genus_Strain)) %>% 
  distinct()
```

```{r}
relabund_random_overlap_tol <- phyl_random_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_tolerant_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  ggtitle(paste("Random: Strains overlapping with Tolerant")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y")
relabund_tolerant_overlap_rand <- phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_random_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  ggtitle(paste("Tolerant: Strains overlapping with Random")) +
  ggthemes::scale_color_few() +
  facet_wrap(~Genus_Strain, scales = "free_y")
```

```{r}
rand_tol_relabund_overlap_plot <-
  phyl_random_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_tolerant_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
         Syncom = "Random") %>% 
  rbind(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_random_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
         Syncom = "Tolerant") ) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  ggtitle(paste("Relative abundance of overlapping isolates")) +
  ggthemes::scale_color_few() +
  facet_wrap(~paste(Syncom, Genus_Strain, sep = "\n"),
             scales = "free_y",
             ncol = 5,
             strip.position = "top") +
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_blank())
```

```{r}
relabund_random_overlap_tol /
  relabund_tolerant_overlap_rand +
  plot_layout(guides = "collect") &
    theme(text = element_text(family ="NimbusMon"),
        strip.background = element_rect(fill = "white"),
        legend.position = "bottom")
```

## Fig 4

### Recovered isolates

```{r}
tolerant_strains_genus <- phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Tolerant") %>% 
  distinct()
random_strains_genus <- phyl_random_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Random") %>% 
  distinct()
sensitive_strains_genus <- phyl_sensitive_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Sensitive") %>% 
  distinct()
mixed_strains_genus <- phyl_mixed_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Mixed") %>% 
  distinct()
```

```{r}
rbind(tolerant_strains_genus, random_strains_genus, sensitive_strains_genus, mixed_strains_genus) %>% 
  write_csv("Found_isolates.csv")
```

### Plot of overlapping strains

```{r}
rand_tol_relabund_overlap_plot <-
  phyl_random_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_tolerant_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
         Syncom = "Random") %>% 
  rbind(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_random_rel_strain %>% 
  psmelt() %>% .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
         Syncom = "Tolerant") ) %>% 
  ggplot(aes(y = Abundance, x = Timepoint, color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(aes(group = Treatment), fun = "mean", geom = "line") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 4) +
  ggtitle(paste("Relative abundance of overlapping isolates")) +
  ggthemes::scale_color_few() +
  facet_wrap(paste(Syncom, Genus_Strain, sep = "\n")~.,
             scales = "free_y",
             ncol = 5,
             strip.position = "top") +
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_blank())
```

## Abundances Random vs tolerant

```{r}
abundances_tol_rand <- bind_rows(list(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain, Treatment, Timepoint, Abundance) %>% 
  mutate(Syncom = "Tolerant"),
  phyl_random_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain, Treatment, Timepoint, Abundance) %>% 
  mutate(Syncom = "Random")))
abundances_tol_rand_summarized <- abundances_tol_rand %>% 
  group_by(Syncom, Genus, Strain, Treatment) %>% 
  summarize(mean = mean(Abundance, na.rm = T),
            sd = sd(Abundance, na.rm = T)) %>%
  ungroup
```

```{r}
abundances_tol_rand_summarized %>% filter(mean > 0.01) %>% 
  ggplot() +
  geom_jitter(aes(y = mean, x = Syncom, color = Genus)) +
  facet_wrap(~Treatment) +
  theme_bw()
```
### Tables

```{r}
  left_join(abundances_tol_rand_summarized %>%
  filter(Syncom == "Random", Treatment %in% c("BOA","Control")) %>%
  dplyr::select(Genus, mean) %>% 
  group_by(Genus) %>% 
  summarize(Random = mean(mean)*100) %>%
  dplyr::select(Genus, Random),
  abundances_tol_rand_summarized %>%
  filter(Syncom == "Tolerant", Treatment %in% c("BOA","Control")) %>%
  dplyr::select(Genus, mean) %>% 
  group_by(Genus) %>% 
  summarize(Tolerant = mean(mean)*100) %>%
  dplyr::select(Genus, Tolerant)) %>% 
    mutate(Tolerant = case_when(is.na(Tolerant) ~ 0,
                              TRUE ~ Tolerant),
           Random = case_when(is.na(Random) ~ 0,
                              TRUE~Random))  %>% 
  mutate(Diff_CTRL = round((Random - Tolerant),2)) %>% 
  arrange(desc(abs(Diff_CTRL)))
```

```{r}
  left_join(abundances_tol_rand_summarized %>%
  filter(Syncom == "Random", Treatment == "APO") %>%
  mutate(Random = mean*100) %>%
  dplyr::select(Treatment, Genus, Random),
  abundances_tol_rand_summarized %>%
  filter(Syncom == "Tolerant", Treatment == "APO") %>%
  mutate(Tolerant = mean*100) %>%
  dplyr::select(Treatment, Genus, Tolerant)) %>% 
    mutate(Tolerant = case_when(is.na(Tolerant) ~ 0,
                              TRUE ~ Tolerant),
           Random = case_when(is.na(Random) ~ 0,
                              TRUE~Random))  %>% 
  mutate(Diff_APO = (Random - Tolerant)*100) %>% 
  arrange(desc(abs(Diff_APO)))
```

## Figure 4A

```{r}
genus_abundance_plot_random_tolerant <- abundances_tol_rand %>% 
  mutate(Strain_Identity = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Same strain",
                               Strain %in% tol_strains ~ "Different strains",
                               Strain %in% rand_strains ~ "Different strains",
                               TRUE ~"Neither")) %>%
  filter(Genus %in% (abundances_tol_rand_summarized %>%
                       filter(mean >= 0.01) %$%
                       Genus),
         !Genus %in% c("Rhizobium", "Hoeflea", "Flavobacterium")
         ) %>% 
  ggplot(aes(y = Abundance,
             x = Syncom, 
             color = fct_relevel(Treatment,"Control","BOA","APO"),
             group = fct_relevel(Treatment,"Control","BOA","APO"),
             shape = fct_relevel(Strain_Identity, "Same strain") )) +
  stat_summary(geom = "point", fun = "mean", position = position_dodge(width = 0.2), size = 4) +
  stat_summary(geom = "errorbar", 
               fun.data = "mean_cl_boot",
               width = 0.2,
               position = position_dodge(width = 0.2),
               color = "black") +
  facet_grid(~Genus) +
  theme_bw() +
  theme(legend.position = "bottom",
        text = element_text(family = "Arial"), 
        strip.background = element_rect(fill = "white"),
        #strip.text.y = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.x = element_blank()) +
  scale_shape_discrete(name = "Strain Identity") +
  scale_color_manual(name =" Treatment",values = rev(ggthemes::few_pal(palette = "Medium")(3)))
genus_abundance_plot_random_tolerant
```

```{r}
genus_log_abundance_plot_random_tolerant <- abundances_tol_rand %>% 
  mutate(Strain_Identity = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Same strain",
                               Strain %in% tol_strains ~ "Different strains",
                               Strain %in% rand_strains ~ "Different strains",
                               TRUE ~"Neither")) %>%
  mutate(Abundance = Abundance + 1e-5) %>% #make log compatible
  ggplot(aes(y = log10(Abundance), x = Genus, color = Syncom, shape = fct_relevel(Strain_Identity, "Same strain") )) +
  stat_summary(geom = "point", fun = "mean", position = position_dodge(width = 0.2), size = 2) +
  stat_summary(aes(group = Syncom),
               geom = "errorbar", 
               fun.data = "mean_cl_boot",
               color = "black",
               width = 0.2,
               position = position_dodge(width = 0.2)) +
  facet_grid(Treatment ~.) +
  theme_bw() +
  theme(legend.position = "bottom",
        text = element_text(family = "Arial"), 
        strip.background = element_rect(fill = "white"),
        #strip.text.y = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.x = element_blank()) +
  scale_shape_discrete(name = "Strain Identity") +
  scale_color_manual(values = syncom_colors) 
```

```{r}
genus_log_abundance_plot_random_tolerant
```

```{r}
genus_low_abundance_plot_random_tolerant
```

```{r}
genus_abundance_plot_random_tolerant_full <- abundances_tol_rand %>% 
  mutate(Strain_Identity = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Same strain",
                               Strain %in% tol_strains ~ "Different strains",
                               Strain %in% rand_strains ~ "Different strains",
                               TRUE ~"Neither")) %>%
  filter(Genus %in% (abundances_tol_rand_summarized %>%
                       filter(mean > 0.01) %$%
                       Genus)) %>% 
  ggplot(aes(y = Abundance, x = Genus, color = Syncom, shape = fct_relevel(Strain_Identity, "Same strain") )) +
  stat_summary(geom = "errorbar", 
               fun.data = "mean_cl_boot",
               width = 0.2,
               position = position_dodge(width = 0.2)) +
  stat_summary(geom = "point", fun = "mean", position = position_dodge(width = 0.2), size = 2) +
  facet_grid(Treatment ~.) +
  theme_bw() +
  theme(legend.position = "bottom",
        text = element_text(family = "Arial"), 
        strip.background = element_rect(fill = "white"),
        #strip.text.y = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.x = element_blank()) +
  scale_shape_discrete(name = "Strain Identity") +
  scale_color_manual(values = syncom_colors) 
```

```{r}
genus_abundance_plot_random_tolerant_full
```

## Figure 4B

### Partial Correlations

```{r}
corr_meth <- "pearson"
adjust <- "fdr"

tol_apo_others_partial <- bind_rows(list(
  phyl_tolerant_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 "APO",
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>%
    mutate(Treatment = "APO",
           Syncom = "Tolerant"),
phyl_tolerant_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 c("BOA", "DMSO"),
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>% 
  mutate(Treatment = "no APO",
          Syncom = "Tolerant")))

rand_apo_others_partial <- bind_rows(list(
  phyl_random_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 "APO",
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>%
    mutate(Treatment = "APO",
           Syncom = "Random") ,
  phyl_random_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 c("BOA", "DMSO"),
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>% 
  mutate(Treatment = "no APO",
         Syncom = "Random"))) 

tol_rand_apo_others_partial <- rbind(tol_apo_others_partial, rand_apo_others_partial)
tol_rand_apo_others_p_partial <- tol_rand_apo_others_partial %>% 
  filter(p < 0.1)
```

```{r}
abundances_tol_rand_summarized_apo_others <- abundances_tol_rand %>% 
  mutate(Treatment = case_when(Treatment == "APO" ~ Treatment,
                               TRUE ~"no APO")) %>% 
  group_by(Syncom, Genus, Strain, Treatment) %>% 
  summarize(mean = mean(Abundance, na.rm = T),
            sd = sd(Abundance, na.rm = T)) %>%
  ungroup
```

```{r}
set.seed(5)
plot_partial_corrs_dat_apo_others <- tol_rand_apo_others_p_partial %>% 
  igraph::graph_from_data_frame() %>% 
  ggnetwork::ggnetwork(layout = igraph::with_fr(),
                       arrow.gap = F) %>% 
  distinct() %>% 
  fix_node_placement_no_time(Treatments = list("no APO","APO")) %>% 
  mutate(Genus = name) %>%  
  left_join(rbind(tolerant_strains_genus, random_strains_genus),
            by = c("Syncom","Genus")) %>% 
  mutate(presence = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Both",
                               Strain %in% tol_strains ~ "Only one",
                               Strain %in% rand_strains ~ "Only one",
                               TRUE ~"Neither")) %>% 
  left_join(abundances_tol_rand_summarized_apo_others, by = c("Syncom","Genus","Strain","Treatment"))
```


```{r}
plot_partial_corrs_apo_others <- plot_partial_corrs_dat_apo_others %>% 
  mutate(x = case_when(x == 1 ~ 0.7, TRUE ~ x ),
           xend = case_when(xend == 1 ~ 0.7, TRUE ~ xend)) %>% 
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(color = r), size = 1.5, curvature = 0.2, alpha = 0.9) +
  geom_nodes(aes(size = mean + sd), color = "black") + 
  geom_nodes(aes(size = mean - sd), color = "white") + 
  geom_nodelabel_repel(aes(label = paste(Genus, Strain, sep = "\n")),  seed = 342, size= 3) +
  geom_nodelabel_repel(aes(label = paste(Genus, Strain, sep = "\n"),  fill = presence), alpha = 0.3, seed=342, size= 3) +
 # ggtitle("Partial correlations") +
  scale_color_viridis_c(option = "C") +
  ggthemes::scale_fill_canva(palette = 'Corporate and sleek') +
  facet_grid(Treatment~Syncom,
             scales = "free"
             ) +
  theme_blank() +
  theme(text = element_text(family = "Arial"),
        strip.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA)) +
  # lims(x = c(-0.1,1.1),
  #      y = c(-0.1,1.1)) +
  NULL
```
```{r}
genus_abundance_plot_random_tolerant / plot_partial_corrs_apo_others +
  plot_layout(nrow = 2,heights = c(1,2), guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(legend.position = "bottom",
        text = element_text(family = "Arial")) 
```
```{r}
plot_partial_corrs_apo_others
```



```{r}
rand_tol_relabund_overlap_plot_avg_timepoints <-
  phyl_random_rel_strain %>% 
  psmelt() %>%
  filter(Strain %in% (phyl_tolerant_rel_strain %>% 
                      psmelt() %>% 
                      .$Strain)) %>% 
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
         Syncom = "Random") %>% 
  rbind(phyl_tolerant_rel_strain %>% 
    psmelt() %>%
    filter(Strain %in% (phyl_random_rel_strain %>% 
                                    psmelt() %>%
                                    .$Strain)) %>% 
    mutate(Genus_Strain = paste(Genus, Strain, sep = "\n"),
           Syncom = "Tolerant") ) %>% 
  ggplot(aes(y = Abundance,
             x = fct_relevel(Treatment, c("Control")), 
             color = Treatment)) +
  stat_summary(aes(group = Treatment), fun.data = "mean_cl_boot", geom = "pointrange") +
  theme_bw() + 
  scale_y_continuous(n.breaks = 3) +
  labs(y = "Relative abundance") +
  ggthemes::scale_color_few() +
  facet_wrap(paste(Syncom, Genus_Strain, sep = "\n") ~.,
              scales = "free_y",
              ncol = 2,
              dir = "v",
              strip.position = "top") +
  #facet_grid(Genus_Strain ~ Syncom) +
  theme(legend.position = "bottom",
        text = element_text(family = "Arial"), 
        strip.background = element_rect(fill = "white"),
        strip.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())
```

```{r}
genus_abundance_plot_random_tolerant +
  plot_partial_corrs + 
  plot_layout(ncol = 2,widths = c(1,2), guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(legend.position = "bottom",
        text = element_text(family = "Arial")) 
```



