---
title: "Scripts for figures for Plant-derived benzoxazinoids act as antibiotics and shape bacterial communities"
author: "Niklas Schandry"
date: "27 4 2021"
output:
  pdf_document:
    toc: true
---

# About

This file contains code for the generation of the main Figures in Schandry et al. Plant-derived benzoxazinoids act as antibiotics and shape bacterial communities.
Detailed scripts for the analysis of individuals and trees can be found in _Growth_trees.Rmd_. The read processing is documented in _Reads_in.Rmd_ and the community analysis is documented in _Community_analysis.Rmd_.
\newpage

# Prepare session

## Libraries and functions

```{r setup, message = F}
library(tidyverse)
library(phyloseq)
library(patchwork)
library(magrittr)
library(ggnetwork)
library(correlation)
library(ggtree)
library(treeio)

source("functions/get_counts_and_overlay.R")
source("functions/phyloseq_corrs.R")
source("functions/overlay_corrnet.R")
source("functions/fix_node_placement.R")
source("functions/compare_conditions.R")
source("functions/read_fasta.R")
```

```{r}
syncom_colors <- c("Random" = "#666699",
                  "Tolerant" = "#9BCB40",
                  "Mixed" = "#B6539F",
                  "Sensitive" = "#FEE377")
```

# Figure 1

## Data in

```{r}
### Taxonomy
tax_info <- read_rds("files_publication/taxonomy.rds")


### read tree

fit_GTR <- read_rds("files_publication/Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")

### read growth data
AUC_emmeans <- read_rds("files_publication/AUC_emmeans_pub.rds") %>%
  mutate(Family = case_when(Genus == "Rhizobacter" ~ "Burkholderiales",
                            TRUE ~ Family)) %>% 
  left_join(tax_info %>%
            dplyr::select(Genus, Class, Order), by = "Genus") %>% 
  filter(Strain %in% fit_GTR$tree$tip.label) ## Keep only those with sequences

tree_data <- AUC_emmeans %>% 
  mutate(id = str_c(Genus,Strain ,sep ="_"))

fit_GTR$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname

syncoms <-  rbind(
                  read_rds("files_publication/syncom_reduced_mixed.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Mixed"),
                 read_rds("files_publication/syncom_reduced_tolerant.rds") %>%
                   dplyr::select(Strain) %>% 
                   mutate(Syncom = "Reduced Tolerant"),
                 read_rds("files_publication/syncom_reduced_sensitive.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Sensitive"),
                 read_rds("files_publication/syncom_full_random.rds") %>%
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Random")
                 )
# Define colors for plotting
set.seed(42069)
family_colors <-  sample(c('#543005','#8c510a','#bf812d',
                           '#dfc27d','#f6e8c3','#4d9221',
                           '#c7eae5','#80cdc1','#35978f',
                           '#01665e','#003c30','#8e0152',
                           '#c51b7d','#de77ae','#f1b6da',
                           '#fde0ef','#000000','#e6f5d0',
                           '#b8e186','#7fbc41','#fb8072',
                           '#276419','#40004b','#bababa'),24)
names(family_colors) <- AUC_emmeans$Family %>% unique %>% sort
```

## Figure 1A

### Prepare data

```{r}
### Below creates "Group Info", which is currently the Genus. Extracted from tiplabels, see above chunk for modification of tiplabels.
tree_structure <- fit_GTR$tree %>% as.phylo
groupInfo <- split(fit_GTR$tree$tip.label, gsub("_\\w+", "", fit_GTR$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)

tree_plot_data <- cbind(
           tree_data %>%
             filter(Chem == "APO") %>% 
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id, Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             as.data.frame() %>% 
             dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
             separate(id, into = c("Genus","Strain"), sep = "_", remove = FALSE) %>% 
             left_join(., AUC_emmeans %>% dplyr::select(Strain, Family, Order, Class, Phylum), by = "Strain") %>% 
             distinct()  ,
          dummy1 = c(rep(NA,174)),
          dummy2 = c(rep(NA,174)),
         tree_data %>% filter(Chem == "BOA")  %>%
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id,Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             dplyr::select("BOA10","BOA50","BOA100") %>% 
             as.data.frame()) 



tree_data_boa <-  tree_plot_data %>% 
           dplyr::select("id","BOA10","BOA50","BOA100") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("10µM BOA", "50µM BOA", "100µM BOA"))
tree_data_apo <- tree_plot_data %>% 
           dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("1µM APO", "5µM APO", "10µM APO", "50µM APO"))
# Make a wide table of memberships, since syncoms were rbinded, there is one entry for each strain/syncom combination.


syncom_memberships <- left_join(tree_plot_data %>% dplyr::select(Strain, id),
  syncoms %>%
    mutate(member = as.factor(case_when(
                              Syncom == "Reduced Mixed" ~ "Mixed",
                              Syncom == "Reduced Tolerant" ~ "Tolerant",
                              Syncom == "Reduced Sensitive" ~ "Sensitive",
                              Syncom == "Random" ~ "Random",
                              TRUE ~ "NA"))) %>%
             pivot_wider(names_from = Syncom, values_from = member),
  by = "Strain")
```

### Figure 1A -  Plot

```{r}
fan_tree_family <- tree_structure %>%
  ggtree(branch.length = "none",
         layout = "fan",
         open.angle = 10) %<+% (tree_plot_data %>% mutate(Phylum_Family = paste(Phylum, Family, sep = "_")))

fan_tree_family <- fan_tree_family +
    geom_tippoint(aes(color =  Family,
                      shape = Class),
                  size = 1) +
  scale_color_manual("Family",
                     values = family_colors,
                     guide = "legend") +
  scale_shape(breaks = c("Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli"))
```

```{r}
fan_tree_family <- fan_tree_family + guides(Family = guide_legend(ncol = 3, byrow = TRUE))
heat_boa_fam <- gheatmap(fan_tree_family,
         tree_data_boa,
         offset=0.3,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth")  +
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth")
heat_boa_apo_fam <- gheatmap(heat_boa_fam, 
         tree_data_apo,
         offset=5.7,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon"))
heat_boa_apo_fam <- heat_boa_apo_fam + ggnewscale::new_scale_fill()
heat_boa_apo_fam_syncom <- gheatmap(heat_boa_apo_fam, 
         syncom_memberships %>% 
           dplyr::select("id","Random",
                              "Reduced Tolerant",
                              "Reduced Mixed",
                              "Reduced Sensitive") %>% 
           column_to_rownames("id"),
         offset=11.5,
         width=0.2,
         colnames_angle = 90,
         hjust = 0.5,
         font.size = 0) + 
  scale_fill_manual(na.value = "grey80",
                           name = "Syncom",
                           values = syncom_colors) +
  theme(text = element_text(family = "NimbusMon"),
        legend.position = "left") 
heat_boa_apo_fam_syncom + plot_annotation(title = "Fig 1") & theme(text = element_text(family = "NimbusMon")) 
```

## Figure 1B - Plot

```{r}
AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO") %>%
  ggplot() +
  geom_point(aes(y = Family, 
                 x = estimate, 
                 fill = Family,
                 group = Strain),
             pch =21,
             alpha=0.3,
             size = 3,
             stat = "summary") +
  ggridges::geom_density_ridges(aes(x=AUC_norm, 
                                    y=Family,
                   color = Family,
                   fill = Family), 
                   alpha = 0.4) +
  facet_wrap(~fct_relevel(Class,
                          "Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli"),
             scales = "free_y") +
  scale_color_manual(values = family_colors) +
  scale_fill_manual(values = family_colors) +
  theme_bw(base_family = "NimbusMon") +
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "AUC[APO] / AUC[Control]") +
  lims(x=c(-0.3,1.6)) +
  NULL
```
\newpage

# Figure 2

## Data in

Read the sequence tables etc (see _Reads_in.Rmd_)

### Random

```{r read objects for phyloseq random}
# Read sequence table
syncomRandom_noChim    <-  readRDS("files_publication/syncom_full_random_seqtable_nochim.rds")
syncomRandom_tax       <-  dada2::assignTaxonomy(seqs = syncomRandom_noChim,
                                    "files_publication/syncom_full_random_16SrRNA.fasta",
                                     taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomRandom_metadata  <-  readRDS("files_publication/Metadata_publication.rds") %>% 
                              filter(Syncom == "Full_Random")

syncomRandom_strains <-  read_rds("files_publication/syncom_full_random.rds") %>%
  mutate(SynCom = "Random")
```

```{r,message = FALSE}
phyl_random <- phyloseq(otu_table(syncomRandom_noChim, taxa_are_rows =F),
                  sample_data(syncomRandom_metadata),
                  tax_table(syncomRandom_tax),
                  phy_tree(read_rds("files_publication/syncom_full_random_GTR_tree.rds")$tree))

phyl_random_glom_strain <- phyl_random %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_random_rel_strain <- transform_sample_counts(phyl_random_glom_strain, function(x) x / sum(x) ) 
```

### Tolerant

```{r read objects for phyloseq tolerant}
# Read sequence table
syncomTolerant_noChim   <-  readRDS("files_publication/syncom_reduced_tolerant_seqtable_nochim.rds")

syncomTolerant_tax      <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_tolerant_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomTolerant_metadata <-  readRDS("files_publication/Metadata_publication.rds") %>%
                          filter(Syncom == "Reduced_Tolerant")

syncomTolerant_strains <-  read_rds("files_publication/syncom_reduced_tolerant.rds") %>%
  mutate(SynCom = "Tolerant")
```

```{r,message = FALSE}
phyl_tolerant <- phyloseq(otu_table(syncomTolerant_noChim, taxa_are_rows =F),
                  sample_data(syncomTolerant_metadata),
                  tax_table(syncomTolerant_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_tolerant_GTR_tree.rds")$tree))

phyl_tolerant_glom_strain <- phyl_tolerant %>% 
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_tolerant_rel_strain <- transform_sample_counts(phyl_tolerant_glom_strain, function(x) x / sum(x) ) 
```

### Mixed


```{r read objects for phyloseq mixed}
# Read sequence table
syncomMixed_noChim <-  readRDS("files_publication/syncom_reduced_mixed_seqtable_nochim.rds")
syncomMixed_tax <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_mixed_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomMixed_metadata        <-  readRDS("files_publication/Metadata_publication.rds") %>%                                   filter(Syncom == "Reduced_Mixed")

syncomMixed_strains <-  read_rds("files_publication/syncom_reduced_mixed.rds") %>%
  mutate(SynCom = "Mixed")
```

```{r,message = FALSE}
phyl_mixed <- phyloseq(otu_table(syncomMixed_noChim, taxa_are_rows =F),
                  sample_data(syncomMixed_metadata),
                  tax_table(syncomMixed_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_mixed_GTR_tree.rds")$tree))

phyl_mixed_glom_strain <- phyl_mixed %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_mixed_rel_strain <- transform_sample_counts(phyl_mixed_glom_strain, function(x) x / sum(x) ) 
```

### Sensitive


```{r read objects for phyloseq sensitive}
# Read sequence table
syncomSensitive_noChim <-  readRDS("files_publication/syncom_reduced_sensitive_seqtable_nochim.rds")
syncomSensitive_tax <- dada2::assignTaxonomy(seqs = syncomTolerant_noChim,
                                     "files_publication/syncom_reduced_sensitive_16SrRNA.fasta",
                                      taxLevels = c("Kingdom", "Phylum", "Family", "Genus", "Strain"))
syncomSensitive_metadata        <-  readRDS("files_publication/Metadata_publication.rds") %>%                                   filter(Syncom == "Reduced_Sensitive")

syncomSensitive_strains <-  read_rds("files_publication/syncom_reduced_sensitive.rds") %>%
  mutate(SynCom = "Tolerant")
```

```{r,message = FALSE}
phyl_sensitive <- phyloseq(otu_table(syncomSensitive_noChim, taxa_are_rows =F),
                  sample_data(syncomSensitive_metadata),
                  tax_table(syncomSensitive_tax),
                  phy_tree(read_rds("files_publication/syncom_reduced_sensitive_GTR_tree.rds")$tree))

phyl_sensitive_glom_strain <- phyl_sensitive %>%
  subset_samples(Rep != "Rep_1") %>%
  tax_glom("Strain")
phyl_sensitive_rel_strain <- transform_sample_counts(phyl_sensitive_glom_strain, function(x) x / sum(x) ) 
```


## Plot

### Individual panels

```{r}
cca_jaccard_random_rel_strain <-  phyl_random_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_random_plot<- plot_ordination(phyl_random_rel_strain,
                cca_jaccard_random_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Random Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_tolerant_rel_strain <-  phyl_tolerant_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_tolerant_plot <- plot_ordination(phyl_tolerant_rel_strain,
                cca_jaccard_tolerant_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Tolerant Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_mixed_rel_strain <-  phyl_mixed_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_mixed_plot <- plot_ordination(phyl_mixed_rel_strain,
                cca_jaccard_mixed_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Mixed Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  

cca_jaccard_sensitive_rel_strain <-  phyl_sensitive_rel_strain %>%
  ordinate("CCA", "jaccard", formula = ~Treatment+Timepoint) 
cca_jaccard_sensitve_plot <- plot_ordination(phyl_sensitive_rel_strain,
                cca_jaccard_sensitive_rel_strain,
                color = "Treatment",
                axes = c(1,2)) +
  ggforce::geom_mark_ellipse(fill = "grey90") +
  geom_point() +
  facet_grid(~Timepoint) +
  geom_point(size = 3) +
  geom_point(size = 3, pch = 21, color = "black") +
  ggtitle(paste("Sensitive Syncom: CCA")) +
  ggthemes::scale_color_few() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",face = "bold"),
        strip.background = element_rect(fill = "white"))  
```

### Figure

```{r}
(cca_jaccard_random_plot | cca_jaccard_tolerant_plot) /
(cca_jaccard_mixed_plot | cca_jaccard_sensitve_plot) +
  plot_layout(guides = "collect",
              heights = c(1,1)) +
  plot_annotation(tag_levels = "A", caption = "Formula: Treatment + Timepoint") &
  theme(text=element_text(family = "NimbusMon"), strip.background = element_rect(fill = "white"))
```
\newpage

# Figure 3

## Calculate l2fc

```{r}
all_lf2cs_no_time <- bind_rows(list(
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Mixed", Treatment = "APO"),
phyl_mixed_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Mixed", Treatment = "BOA"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO", 
                     treat2 = "Control", 
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Sensitive", Treatment = "APO"),
phyl_sensitive_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Sensitive", Treatment = "BOA"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05, 
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Random", Treatment = "APO"),
phyl_random_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>% 
  mutate(Syncom = "Random", Treatment = "BOA"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "APO",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F, 
                     return_unfiltered = T) %>%
  mutate(Syncom = "Tolerant", Treatment = "APO"),
phyl_tolerant_glom_strain %>% 
  compare_conditions(treatvar = "Treatment",
                     treat1 = "BOA",
                     treat2 = "Control",
                     alpha = 0.05,
                     OTUcol = "Strain",
                     plot = F,
                     return_unfiltered = T) %>%
  mutate(Syncom = "Tolerant", Treatment = "BOA")))
```

```{r}
strain_AUC <- AUC_emmeans %>%
  filter(Treat.x %in% c("Control","50uM_APO","100uM_BOA")) %>%
  mutate(Treatment = case_when(Treat.x == "Control" ~ Treat.x,
                               Treat.x == "50uM_APO" ~ "APO",
                               Treat.x == "100uM_BOA" ~ "BOA")) %>% 
  dplyr::select(Strain, Genus ,Treatment, AUC, estimate) %>%
  group_by(Strain,Treatment) %>%
  summarize(mean_AUC = mean(AUC, na.rm = T),
            mean_rel_AUC = mean(estimate, na.rm = T)) %>%
  ungroup  
```

## Plot l2FC community vs AUC ratio

```{r}
growth_data_l2fc <- left_join(strain_AUC, all_lf2cs_no_time,  by = c("Strain", "Treatment")) %>% 
  filter(!is.na(Treatment),
         !is.na(Syncom),
         !is.na(log2FoldChange),
         Treatment != "Control") 
```


## Figure 3A - Plot

```{r}
growth_data_l2fc %>% 
  ggplot(aes(x = mean_rel_AUC,
             y = log2FoldChange)) +
  geom_vline(aes(xintercept = 1), color = "darkgrey") +
  geom_hline(aes(yintercept = 0), color = "darkgrey") +
  geom_smooth(method = "lm", color = "black", se = T) +
  geom_point(colour = "black", 
             size = 5, 
             alpha = 0.7) +
  geom_point(aes(color = Syncom),
             size = 5, 
             alpha = 0.7) +
  scale_color_manual(values = syncom_colors) +
  facet_wrap(~Treatment, scales = "free_x") +
  theme_bw(base_family = "NimbusMon") +
  xlab(paste("Effect on strain grown in isolation \n [Ratio area under the curve (Treatment / Control)]")) +
  ylab(paste("Effect on strain grown in community \n [log2Foldchange (Treatment / Control)]")) +
  ggtitle("Growth and abundance changes by treatment") +
  theme(plot.title = element_text(family = "NimbusMon"),
        plot.subtitle =  element_text(family = "NimbusMon"),
        text = element_text(family = "NimbusMon"),
        strip.background = element_blank(),
        axis.title = element_text(family = "NimbusMon"),
        strip.text = element_text(family = "NimbusMon",
                                  face = "bold"),
        legend.position = "none") 
```

## Figure 3B - Plot

```{r}
all_lf2cs_no_time %>% 
  left_join(strain_AUC, by = c("Strain", "Treatment")) %>% 
  mutate(genus_strain = str_c(Genus,Strain, sep = "\n")) %>% 
  group_by(Strain) %>% 
  add_count(Treatment) %>% 
  filter(n > 1) %>% 
  ggplot(aes( y=log2FoldChange, x = Treatment, group = Syncom)) +
  facet_wrap(~genus_strain, ncol = 5, scales = "free_x") +
  geom_hline(aes(yintercept = 0),
             linetype = "dashed",
             color = "darkgrey" ) +
  geom_errorbar(aes(ymin = log2FoldChange-lfcSE, 
                      ymax = log2FoldChange + lfcSE),
                 position = position_dodge(width = 0.6),
                 width = 0.2) +
  geom_point(size = 4,
             position = position_dodge(width = 0.6)) +
  geom_point(aes(color = Syncom),
             size = 3,
             position = position_dodge(width = 0.6)) +
  geom_vline(aes(xintercept = 1.5)) +
  scale_x_discrete(limits = rev) + 
  coord_flip() +
  theme_bw() +
  theme(text = element_text(family = "NimbusMon"),
        strip.background = element_rect(fill = "white"),
        legend.position = "none") +
  scale_color_manual(values = syncom_colors) +
  labs(title = "Changes in abundance",
       sub = "Strains included in more than one community",
       y = "log2FC vs Control (across all timepoints)",
       x = "Treatment")
```
\newpage

# Figure 4

## Figure 4A

### Abundance table

```{r}
abundances_tol_rand <- bind_rows(list(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain, Treatment, Timepoint, Abundance) %>% 
  mutate(Syncom = "Tolerant"),
  phyl_random_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain, Treatment, Timepoint, Abundance) %>% 
  mutate(Syncom = "Random")))
abundances_tol_rand_summarized <- abundances_tol_rand %>% 
  group_by(Syncom, Genus, Strain, Treatment) %>% 
  summarize(mean = mean(Abundance, na.rm = T),
            sd = sd(Abundance, na.rm = T)) %>%
  ungroup
```

### Overlap Random and Tolerant

```{r}
tol_strains <- phyl_tolerant_rel_strain %>% 
  psmelt() %>% .$Strain
rand_strains <- phyl_random_rel_strain %>% 
  psmelt() %>% .$Strain
rand_tol_genusstrain <- rbind(phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  dplyr::select(Strain, Genus_Strain),
  phyl_random_rel_strain %>% 
  psmelt() %>%
  mutate(Genus_Strain = paste(Genus, Strain, sep = "\n")) %>% 
  dplyr::select(Strain, Genus_Strain)) %>% 
  distinct()
tolerant_strains_genus <- phyl_tolerant_rel_strain %>% 
  psmelt() %>%
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Tolerant") %>% 
  distinct()
random_strains_genus <- phyl_random_rel_strain %>% 
  psmelt() %>% 
  dplyr::select(Genus, Strain) %>%
  mutate(Syncom = "Random") %>% 
  distinct()
```

### Figure 4A- Plot

```{r}
genus_abundance_plot_random_tolerant <- abundances_tol_rand %>% 
  mutate(Strain_Identity = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Same strain",
                               Strain %in% tol_strains ~ "Different strains",
                               Strain %in% rand_strains ~ "Different strains",
                               TRUE ~"Neither")) %>%
  filter(Genus %in% (abundances_tol_rand_summarized %>%
                       filter(mean >= 0.01) %$%
                       Genus),
         !Genus %in% c("Rhizobium", "Hoeflea", "Flavobacterium")
         ) %>% 
  ggplot(aes(y = Abundance,
             x = Syncom, 
             color = fct_relevel(Treatment,"Control","BOA","APO"),
             group = fct_relevel(Treatment,"Control","BOA","APO"),
             shape = fct_relevel(Strain_Identity, "Same strain") )) +
  stat_summary(geom = "point", fun = "mean", position = position_dodge(width = 0.2), size = 4) +
  stat_summary(geom = "errorbar", 
               fun.data = "mean_cl_boot",
               width = 0.2,
               position = position_dodge(width = 0.2),
               color = "black") +
  facet_grid(~Genus) +
  theme_bw() +
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon"), 
        strip.background = element_rect(fill = "white"),
        #strip.text.y = element_blank(),
        axis.text.x = element_text(angle = 30, hjust = 1),
        axis.title.x = element_blank()) +
  scale_shape_discrete(name = "Strain Identity") +
  scale_color_manual(name =" Treatment",values = rev(ggthemes::few_pal(palette = "Medium")(3)))
genus_abundance_plot_random_tolerant
```

## Figure 4B

## Figure 4B

### Partial Correlations

```{r}
corr_meth <- "pearson"
adjust <- "fdr"

tol_apo_others_partial <- bind_rows(list(
  phyl_tolerant_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 "APO",
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>%
    mutate(Treatment = "APO",
           Syncom = "Tolerant"),
phyl_tolerant_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 c("BOA", "DMSO"),
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>% 
  mutate(Treatment = "no APO",
          Syncom = "Tolerant")))

rand_apo_others_partial <- bind_rows(list(
  phyl_random_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 "APO",
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>%
    mutate(Treatment = "APO",
           Syncom = "Random") ,
  phyl_random_glom_strain %>% 
  phyloseq_corrs("Treatment",
                 c("BOA", "DMSO"),
                 p.adj = adjust,
                 mincounts = 100, 
                 method = corr_meth,
                 OTUcol = "Genus",
                 partial = TRUE) %>% 
  mutate(Treatment = "no APO",
         Syncom = "Random"))) 

tol_rand_apo_others_partial <- rbind(tol_apo_others_partial, rand_apo_others_partial)
tol_rand_apo_others_p_partial <- tol_rand_apo_others_partial %>% 
  filter(p < 0.1)
```

### Node information

```{r}
abundances_tol_rand_summarized_apo_others <- abundances_tol_rand %>% 
  mutate(Treatment = case_when(Treatment == "APO" ~ Treatment,
                               TRUE ~"no APO")) %>% 
  group_by(Syncom, Genus, Strain, Treatment) %>% 
  summarize(mean = mean(Abundance, na.rm = T),
            sd = sd(Abundance, na.rm = T)) %>%
  ungroup
```

### Figure 4B - Plot

```{r}
set.seed(5)
plot_partial_corrs_dat_apo_others <- tol_rand_apo_others_p_partial %>% 
  igraph::graph_from_data_frame() %>% 
  ggnetwork::ggnetwork(layout = igraph::with_fr(),
                       arrow.gap = F) %>% 
  distinct() %>% 
  fix_node_placement_no_time(Treatments = list("no APO","APO")) %>% 
  mutate(Genus = name) %>%  
  left_join(rbind(tolerant_strains_genus, random_strains_genus),
            by = c("Syncom","Genus")) %>% 
  mutate(presence = case_when(Strain %in% intersect(tol_strains, rand_strains) ~ "Both",
                               Strain %in% tol_strains ~ "Only one",
                               Strain %in% rand_strains ~ "Only one",
                               TRUE ~"Neither")) %>% 
  left_join(abundances_tol_rand_summarized_apo_others, by = c("Syncom","Genus","Strain","Treatment"))
```


```{r}
plot_partial_corrs_apo_others <- plot_partial_corrs_dat_apo_others %>% 
  mutate(x = case_when(x == 1 ~ 0.7, TRUE ~ x ),
           xend = case_when(xend == 1 ~ 0.7, TRUE ~ xend)) %>% 
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(color = r), size = 1.5, curvature = 0.2, alpha = 0.9) +
  geom_nodes(aes(size = mean + sd), color = "black") + 
  geom_nodes(aes(size = mean - sd), color = "white") + 
  geom_nodelabel_repel(aes(label = paste(Genus, Strain, sep = "\n")),  seed = 342, size= 3) +
  geom_nodelabel_repel(aes(label = paste(Genus, Strain, sep = "\n"),  fill = presence), alpha = 0.3, seed=342, size= 3) +
 # ggtitle("Partial correlations") +
  scale_color_viridis_c(option = "C") +
  ggthemes::scale_fill_canva(palette = 'Corporate and sleek') +
  facet_grid(Treatment~Syncom,
             scales = "free"
             ) +
  theme_blank() +
  theme(text = element_text(family = "NimbusMon"),
        strip.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA)) +
  # lims(x = c(-0.1,1.1),
  #      y = c(-0.1,1.1)) +
  NULL
```

```{r}
plot_partial_corrs_apo_others
```

## Figure 4 arranged

```{r}
genus_abundance_plot_random_tolerant / plot_partial_corrs_apo_others +
  plot_layout(nrow = 2,heights = c(1,2), guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(legend.position = "bottom",
        text = element_text(family = "NimbusMon")) 
```

