---
title: "Isolate growth"
author: "Niklas Schandry"
date: "2020"
output: 
  html_document:
    fig_caption: yes
## Widescreen figures
    fig_height: 9
    fig_width: 16
## Highlighting    
    highlight: zenburn
    theme: cosmo
### TOC    
    toc: yes
    toc_float: yes
---

# About

This file documents the isolate analysis for Schandry, et al., 2020

# Prepare session

## Load libraries

```{r eval = T, message = F}
library(tidyverse)
library(magrittr)
library(DECIPHER)
library(phangorn)
library(patchwork)
library(ggtree)
require(wesanderson)
source("functions/read_fasta.R")
```

# Compute tree

This tree is saved, chunk not evaluated to save time on knits.

```{r eval = F}
Root_16S <- read_fasta("files_publication/16SrRNA_Root_isolates_WGS.fasta") %>% as.data.frame() %>% rownames_to_column("Strain")
Soil_16S <- read_fasta("files_publication/16SrRNA_Soil_isolates_WGS.fasta") %>% as.data.frame() %>% rownames_to_column("Strain")
all_16s <- bind_rows(Root_16S,Soil_16S) 
all_16s <- all_16s %>% filter(Strain %in% AUC_emmeans$Strain) %>% filter(Strain != "Root763") # Remove strain that lacks Genus information..

colnames(all_16s) <- c("Strain", "rRNA_16S")

relevant_16S <- all_16s %$% DNAStringSet(rRNA_16S)
names(relevant_16S) <- all_16s$Strain



seqs_aligned <- AlignSeqs(relevant_16S, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Strain_tree.rds")
```

# Refined Taxonomic database

For reasons that are not completely clear, the order and class information is missing from the tables.
This information will be added from the bac120_taxonomy.tsv (genome taxonomy database https://www.nature.com/articles/nbt.4229) 

```{r eval = F}
tax_info <- read_tsv("files_publication/bac120_taxonomy.tsv") %>%
  set_colnames(c("col1", "col2")) %>% 
  dplyr::select(col2) %>% 
  mutate(col2 = col2 %>% str_remove_all("[a-z]__")) %>% 
  separate(col2, c("Domain","Phylum","Class","Order","Family","Genus","Species"), sep = ";") %>% 
  dplyr::select(-Species) %>% 
  unique()
# Add missing information
tax_info <- rbind(tax_info,
            c("Bacteria",
              "Proteobacteria",
              "Betaproteobacteria",
              "Burkholderiales",
              "Comamonadaceae",
              "Acidovorax")) %>% 
            rbind(
             c("Bacteria",
               "Proteobacteria",
               "Gammaproteobacteria",
               "Xanthomonadales",
               "Xanthomonadaceae",
               "Pseudoxanthomonas"))

write_rds(tax_info, "data/taxonomy.rds")
```

# Read files

Read the growth data and the tree

```{r}
tax_info <- read_rds("files_publication/taxonomy.rds")
### read growth data
AUC_emmeans <- read_rds("files_publication/AUC_emmeans_pub.rds") %>% 
  left_join(tax_info %>%
            dplyr::select(Genus, Class, Order), by = "Genus")
tree_data <- AUC_emmeans %>% 
  mutate(id = str_c(Genus,Strain ,sep ="_"))

### read tree
fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname

syncoms <-  rbind(
                  read_rds("files_publication/syncom_reduced_mixed.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Mixed"),
                 read_rds("files_publication/syncom_reduced_tolerant.rds") %>%
                   dplyr::select(Strain) %>% 
                   mutate(Syncom = "Reduced Tolerant"),
                 read_rds("files_publication/syncom_reduced_sensitive.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Sensitive"),
                 read_rds("files_publication/syncom_full_random.rds") %>%
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Random")
                 )

```

# Some descriptive stats of the growth data

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("1uM_APO","5uM_APO","10uM_APO","50uM_APO")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("1uM_APO","5uM_APO","10uM_APO","50uM_APO"))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  group_by(Treatment) %>% 
  mutate(Sensitive = case_when(estimate > 0.5 ~ "No",
                                   TRUE ~ "Yes")) %>% 
  group_by(Treatment, Sensitive) %>% 
  tally()
```

## Density plots

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("1uM_APO","5uM_APO","10uM_APO","50uM_APO")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("1uM_APO","5uM_APO","10uM_APO","50uM_APO"))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   #color = Treatment, 
                   fill = Treatment), alpha = 0.3) +
  theme_bw(base_family = "NimbusMon") +
  ggtitle("Estimated relative growth for APO treatments")
```

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("100uM_BOA","50uM_BOA", "10uM_BOA")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("10uM_BOA","50uM_BOA","100uM_BOA" ))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   #color = Treatment, 
                   fill = Treatment), alpha = 0.3) +
  theme_bw(base_family = "NimbusMon") +
    ggtitle("Estimated relative growth for BOA treatments")
```

# Heatmap tree

```{r}
### Below creates "Group Info", which is currently the Genus. Extracted from tiplabels, see above chunk for modification of tiplabels.
tree_structure <- fit_GTR$tree %>% as.phylo
groupInfo <- split(fit_GTR$tree$tip.label, gsub("_\\w+", "", fit_GTR$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
```

```{r}
tree_plot_data <- cbind(
           tree_data %>%
             filter(Chem == "APO") %>% 
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id, Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             as.data.frame() %>% 
             dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
             separate(id, into = c("Genus","Strain"), sep = "_", remove = FALSE) %>% 
             left_join(., AUC_emmeans %>% dplyr::select(Strain, Family, Order, Class, Phylum), by = "Strain") %>% 
             distinct()  ,
          dummy1 = c(rep(NA,180)),
          dummy2 = c(rep(NA,180)),
         tree_data %>% filter(Chem == "BOA")  %>%
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id,Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             dplyr::select("BOA10","BOA50","BOA100") %>% 
             as.data.frame()) 



tree_data_boa <-  tree_plot_data %>% 
           dplyr::select("id","BOA10","BOA50","BOA100") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("10µM BOA", "50µM BOA", "100µM BOA"))
tree_data_apo <- tree_plot_data %>% 
           dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("1µM APO", "5µM APO", "10µM APO", "50µM APO"))
# Make a wide table of memberships, since syncoms were rbinded, there is one entry for each strain/syncom combination.


syncom_memberships <- left_join(tree_plot_data %>% dplyr::select(Strain, id),
  syncoms %>%
    mutate(member = as.factor(case_when(
                              Syncom == "Reduced Mixed" ~ "Mixed",
                              Syncom == "Reduced Tolerant" ~ "Tolerant",
                              Syncom == "Reduced Sensitive" ~ "Sensitive",
                              Syncom == "Random" ~ "Random",
                              TRUE ~ "NA"))) %>%
             pivot_wider(names_from = Syncom, values_from = member),
  by = "Strain")
```

### Family level heatmap

```{r}
fan_tree_family <- tree_structure %>%
  ggtree(branch.length = "none",
         layout = "fan",
         open.angle = 10) %<+% (tree_plot_data %>% mutate(Phylum_Family = paste(Phylum, Family, sep = "_")))
set.seed(42069)
fan_tree_family <- fan_tree_family +
    geom_tippoint(aes(color =  Family,
                      shape = Class),
                  size = 1) +
  scale_color_manual("Family",
                     values = sample(c('#543005','#8c510a','#bf812d',
                                '#dfc27d','#f6e8c3','#4d9221',
                                '#c7eae5','#80cdc1','#35978f',
                                '#01665e','#003c30','#8e0152',
                                '#c51b7d','#de77ae','#f1b6da',
                                '#fde0ef','#000000','#e6f5d0',
                                '#b8e186','#7fbc41','#fb8072',
                                '#276419','#40004b','#bababa'),24),
                     guide = "legend") +
  scale_shape(breaks = c("Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli"))
```

```{r}
fan_tree_family <- fan_tree_family + guides(Family = guide_legend(ncol = 2, byrow = TRUE))
```


```{r}
heat_boa_fam <- gheatmap(fan_tree_family,
         tree_data_boa,
         offset=0.3,
         width=0.2,
         colnames_angle = 90,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth")  +
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth")
heat_boa_apo_fam <- gheatmap(heat_boa_fam, 
         tree_data_apo,
         offset=5.7,
         width=0.2,
         colnames_angle = 90,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon"))
heat_boa_apo_fam
```

### Full family plot

```{r}
fan_tree_family <- fan_tree_family + guides(Family = guide_legend(ncol = 3, byrow = TRUE))
heat_boa_fam <- gheatmap(fan_tree_family,
         tree_data_boa,
         offset=0.3,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth")  +
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth")
heat_boa_apo_fam <- gheatmap(heat_boa_fam, 
         tree_data_apo,
         offset=5.7,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon"))
heat_boa_apo_fam <- heat_boa_apo_fam + ggnewscale::new_scale_fill()
heat_boa_apo_fam_syncom <- gheatmap(heat_boa_apo_fam, 
         syncom_memberships %>% 
           dplyr::select("id","Random",
                              "Reduced Tolerant",
                              "Reduced Mixed",
                              "Reduced Sensitive") %>% 
           column_to_rownames("id"),
         offset=11.5,
         width=0.2,
         colnames_angle = 90,
         hjust = 0.5,
         font.size = 0) + 
  ggthemes::scale_fill_few(na.value = "grey80",
                           name = "Syncom",
                           labels = c("Random",
                                      "Tolerant",
                                      "Mixed",
                                      "Sensitive"),
                           breaks = c("Random",
                                      "Tolerant",
                                      "Mixed",
                                      "Sensitive")) +
  theme(text = element_text(family = "NimbusMon"),
        legend.position = "left") 
heat_boa_apo_fam_syncom + plot_annotation(title = "Fig 1") & theme(text = element_text(family = "NimbusMon")) 
```

```{r}
heat_boa_apo_fam_syncom & theme(legend.position = "none") 
```

### Plot without legend

```{r}
heat_boa_apo_fam_syncom & theme(legend.position = "none") 
```

## Legend only

```{r}
cowplot::ggdraw(cowplot::get_legend(heat_boa_apo_fam_syncom))
```

# Family subset

```{r}
AUC_emmeans %$% Family %>% as.factor() %>% levels
```

```{r}
AUC_emmeans %>% dplyr::select(Family,Genus, Strain) %>% unique() %>% group_by(Family) %>% tally %>% arrange(desc(n), Family)
```

```{r}
AUC_emmeans %>% filter(Family == "Nocardioidaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Family == "Microbacteriaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Phylum == "Firmicutes") %>% dplyr::select(Genus) %>% unique()
```

```{r}
AUC_emmeans %>% colnames
```

## Commamodaceae

```{r}
Comamonadaceae <- AUC_emmeans %>% filter(Family == "Comamonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_comamo <- Comamonadaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Comamonadaceae_tree <- drop.tip(tree_structure, not_comamo) 
```

```{r}
Comamonadaceae_tree <- Comamonadaceae_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
```


```{r}
Comamonadaceae_tree_boa <- gheatmap(Comamonadaceae_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish) +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Comamonadaceae")+
  layout_dendrogram() +
    geom_tippoint(aes(color = Genus)) +
    ggthemes::scale_color_hc()
Comamonadaceae_tree_boa_apo <- gheatmap(Comamonadaceae_tree_boa, 
         tree_data_apo,
         offset=2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish)
```

```{r}
Comamonadaceae_tree_boa_apo
```


## Microbacteriaceae


```{r}
Microbacteria <- AUC_emmeans %>% filter(Family == "Microbacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Microbacteria <- Microbacteria$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Microbacteria_tree <- drop.tip(tree_structure, not_Microbacteria)

Microbacteria_tree <- Microbacteria_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Microbacteria_tree <- Microbacteria_tree + 
  layout_dendrogram() + 
  ggtitle("Microbacteria") +
    geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("Darjeeling2",4,"discrete"))

Microbacteria_tree <- rotate(Microbacteria_tree, 19)
```

```{r}
Microbacteria_tree_boa <- gheatmap(Microbacteria_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Microbacteriaceae, BOA") 


Microbacteria_tree_boa
```

```{r}
Microbacteria_tree_boa_apo <- gheatmap(Microbacteria_tree_boa, 
         tree_data_apo,
         offset=2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Microbacteriaceae") +
  theme(legend.position = "left")
Microbacteria_tree_boa_apo
```


## Rhizobiaceae


```{r}
Rhizobia <- AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_rhizobium <- Rhizobia$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Rhizobia_tree <- drop.tip(tree_structure, not_rhizobium)
```

```{r}
Rhizobia_tree <- Rhizobia_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Rhizobia_tree <- Rhizobia_tree + 
  layout_dendrogram() + 
  ggtitle("Rhizobia") +
    geom_tippoint(aes(color = Genus), size = 2) +
    scale_color_manual(values = wesanderson::wes_palette("BottleRocket2", 2 ,"discrete"))
```


```{r}
Rhizobia_tree_boa <- gheatmap(Rhizobia_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Rhizobiaceae, BOA") 


Rhizobia_tree_boa
```

```{r}
Rhizobia_tree_boa_apo <- gheatmap(Rhizobia_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Rhizobiaceae") +
  theme(legend.position = "left")
Rhizobia_tree_boa_apo
```


```{r}
Rhizobia_tree_apo <- gheatmap(Rhizobia_tree, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Rhizobiaceae, APO")
Rhizobia_tree_apo
```

```{r}
AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% filter(str_detect(Treat.x,"BOA")) %>% dplyr::select(Strain, Conc, estimate) %>% unique %>% filter(Conc == 100) %>% arrange(desc(estimate))
```

Root708 seems to grow particularily well in BOA

```{r}
AUC_emmeans %>% filter(Strain == "Root708") %>% 
  dplyr::select(Treat.x, estimate) %>% unique
```




## Xanthomonadaceae


```{r}
Xantho <- AUC_emmeans %>% filter(Family == "Xanthomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Xantho <- Xantho$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Xantho_tree <- drop.tip(tree_structure, not_Xantho)

Xantho_tree <- Xantho_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Xantho_tree <- Xantho_tree + 
  layout_dendrogram() + 
  ggtitle("Xanthomonadaceae") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Xantho_tree_boa <- gheatmap(Xantho_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Xanthoceae, BOA") 


Xantho_tree_boa
```

```{r}
Xantho_tree_boa_apo <- gheatmap(Xantho_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Xanthomonadaceae") +
  theme(legend.position = "left")
Xantho_tree_boa_apo
```


```{r}
Xantho_tree_apo <- gheatmap(Xantho_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Xanthomonadaceae, APO")
Xantho_tree_apo
```



## Bacillaceae


```{r}
Bacilli <- AUC_emmeans %>% 
  filter(Family == "Bacillaceae") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_Bacilli <- Bacilli$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Bacilli_tree <- drop.tip(tree_structure, not_Bacilli)

Bacilli_tree <- Bacilli_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Bacilli_tree <- Bacilli_tree + 
  layout_dendrogram() + 
  ggtitle("Bacillaceae") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Bacilli_tree_boa <- gheatmap(Bacilli_tree, 
         tree_data_boa,
         offset=0.16,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Bacilliceae, BOA") 


Bacilli_tree_boa
```

```{r}
Bacilli_tree_boa_apo <- gheatmap(Bacilli_tree_boa, 
         tree_data_apo,
         offset = 1.3,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillaceae") +
  theme(legend.position = "left")
Bacilli_tree_boa_apo
```


```{r}
Bacilli_tree_apo <- gheatmap(Bacilli_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillaceae, APO")
Bacilli_tree_apo
```

## Bacillales

All Firmicutes in the dataset are Bacillales (Paenibacillus, Bacillus)

```{r}
Firmi <- AUC_emmeans %>% 
  filter(Phylum == "Firmicutes") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_Firmi <- Firmi$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Firmi_tree <- drop.tip(tree_structure, not_Firmi)

Firmi_tree <- Firmi_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Firmi_tree <- Firmi_tree + 
  layout_dendrogram() + 
  ggtitle("Bacillales") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Firmi_tree_boa <- gheatmap(Firmi_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Bacillales, BOA") 


Firmi_tree_boa
```

```{r}
Firmi_tree_boa_apo <- gheatmap(Firmi_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillales") +
  theme(legend.position = "left")
Firmi_tree_boa_apo
```


```{r}
Firmi_tree_apo <- gheatmap(Firmi_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillales, APO")
Firmi_tree_apo
```


## Nocardioidaceae


```{r}
Nocardio <- AUC_emmeans %>% filter(Family == "Nocardioidaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Nocardio <- Nocardio$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Nocardio_tree <- drop.tip(tree_structure, not_Nocardio)

Nocardio_tree <- Nocardio_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Nocardio_tree <- Nocardio_tree + 
  layout_dendrogram() + 
  ggtitle("Nocardio") +
    geom_tippoint(aes(color = Genus)) +
    scale_color_manual(values = wesanderson::wes_palette("GrandBudapest1", 2 ,"discrete"))
```


```{r}
Nocardio_tree_boa <- gheatmap(Nocardio_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Nocardioceae, BOA") 


Nocardio_tree_boa
```

```{r}
Nocardio_tree_boa_apo <- gheatmap(Nocardio_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Nocardioidaceae") +
  theme(legend.position = "left")
Nocardio_tree_boa_apo
```


```{r}
Nocardio_tree_apo <- gheatmap(Nocardio_tree, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Nocardioceae, APO")
Nocardio_tree_apo
```



## Pseudomonadaceae


```{r}
Pseudom <- AUC_emmeans %>% filter(Family == "Pseudomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Pseudom <- Pseudom$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Pseudom_tree <- drop.tip(tree_structure, not_Pseudom)

Pseudom_tree <- Pseudom_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Pseudom_tree <- Pseudom_tree + 
  layout_dendrogram() + 
  ggtitle("Pseudom") +
    geom_tippoint(aes(color = Genus)) +
    scale_color_manual(values = wesanderson::wes_palette("Moonrise2", 2 ,"discrete"))
```


```{r}
Pseudom_tree_boa <- gheatmap(Pseudom_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames = FALSE,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Pseudomceae, BOA") 


Pseudom_tree_boa
```

```{r}
Pseudom_tree_boa_apo <- gheatmap(Pseudom_tree_boa, 
         tree_data_apo,
         offset=1.4,
         width=0.2,
         colnames = FALSE,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Pseudomonadaceae") +
  theme(legend.position = "left")
Pseudom_tree_boa_apo
```



## Alcaligenaceae

```{r}
Alcali <- AUC_emmeans %>% filter(Family == "Alcaligenaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_alcali <- Alcali$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Alcali_tree <- drop.tip(tree_structure, not_alcali)

Alcali_tree <- Alcali_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Alcali_tree <- Alcali_tree + 
  layout_dendrogram() + 
  ggtitle("Alcaligenaceae") +
    geom_tippoint(aes(color = Genus))
```


```{r}
Alcali_tree_boa <- gheatmap(Alcali_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Alcaligenaceae, BOA") 


Alcali_tree_boa
```

```{r}
Alcali_tree_boa_apo <- gheatmap(Alcali_tree_boa, 
         tree_data_apo,
         offset=0.7,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Alcaligenaceae") +
  theme(legend.position = "left")
Alcali_tree_boa_apo
```


## Flavobacteriaceae


```{r}
Flavobacteriaceae <- AUC_emmeans %>% filter(Family == "Flavobacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_flavo <- Flavobacteriaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Flavobacteriaceae_tree <- drop.tip(tree_structure, not_flavo)
```

```{r}
Flavobacteriaceae_tree <- Flavobacteriaceae_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Flavobacteriaceae_tree <- Flavobacteriaceae_tree + layout_dendrogram()
```


```{r}
Flavobacteriaceae_tree_boa <- gheatmap(Flavobacteriaceae_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Flavobacteriaceae") +
    geom_tippoint(aes(color = Genus)) +
    ggthemes::scale_color_hc()
Flavobacteriaceae_tree_boa_apo <- gheatmap(Flavobacteriaceae_tree_boa, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth")
```
```{r}
Flavobacteriaceae_tree_boa_apo
```

# Family Trees

```{r}
(Rhizobia_tree_boa_apo + Firmi_tree_boa_apo +
  plot_layout(guides = "collect")  &
  theme(legend.position = "left") &
  guides(fill =FALSE)) 
```

```{r}
((Rhizobia_tree_boa_apo + Firmi_tree_boa_apo )/
   (Pseudom_tree_boa_apo + Alcali_tree_boa_apo)) &
  theme(legend.position = "bottom") &
  guides(fill =FALSE)
```


## Less stringent subsets

```{r}
(Rhizobia_tree_boa_apo + Xantho_tree_boa_apo +
   Microbacteria_tree_boa_apo + Firmi_tree_boa_apo) +
  plot_annotation(tag_levels = c("A")) &
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5)) &
  #plot_layout(guides = "collect")  &
#  theme(legend.position = "left") &
  guides(fill =FALSE) 
```

```{r}
(Rhizobia_tree_boa_apo + Xantho_tree_boa_apo + Microbacteria_tree_boa_apo + Firmi_tree_boa_apo +
  plot_layout(guides = "collect")  &
  theme(legend.position = "left") &
  guides(fill =FALSE))
```

# SessionInfo

```{r}
sessionInfo()
```