---
title: "Isolate growth"
author: "Niklas Schandry"
date: "2020"
output: 
  html_document:
    fig_caption: yes
## Widescreen figures
    fig_height: 9
    fig_width: 16
## Highlighting    
    highlight: zenburn
    theme: cosmo
### TOC    
    toc: yes
    toc_float: yes
---

# About

This file documents the isolate analysis for Schandry, et al., 2020

# Prepare session

## Load libraries

```{r setup, eval = T, message = F}
library(tidyverse)
library(magrittr)
library(DECIPHER)
library(phangorn)
library(patchwork)
library(ggtree)
library(wesanderson)
library(motmot)
source("functions/read_fasta.R")
```

# Compute tree

This tree is saved, chunk not evaluated to save time on knits.

```{r eval = F}
Root_16S <- read_fasta("files_publication/16SrRNA_Root_isolates_WGS.fasta") %>% as.data.frame() %>% rownames_to_column("Strain")
Soil_16S <- read_fasta("files_publication/16SrRNA_Soil_isolates_WGS.fasta") %>% as.data.frame() %>% rownames_to_column("Strain")
all_16s <- bind_rows(Root_16S,Soil_16S) 
all_16s <- all_16s %>% filter(Strain %in% AUC_emmeans$Strain) %>% filter(Strain != "Root763") # Remove strain that lacks Genus information..

colnames(all_16s) <- c("Strain", "rRNA_16S")

relevant_16S <- all_16s %$% DNAStringSet(rRNA_16S)
names(relevant_16S) <- all_16s$Strain



seqs_aligned <- AlignSeqs(relevant_16S, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Strain_tree.rds")
```

Make tree of 16S, gyrB, rpoB, recA, atpD, dnaJ and thrC concat

```{r}
Root_16S <- read_fasta("files_publication/16SrRNA_Root_isolates_WGS.fasta") %>% 
  as.data.frame() %>%
    set_colnames("rRNA16S") %>%
  rownames_to_column("Strain")
Soil_16S <- read_fasta("files_publication/16SrRNA_Soil_isolates_WGS.fasta") %>%
  as.data.frame() %>% 
  set_colnames("rRNA16S") %>%
  rownames_to_column("Strain")
all_16s <- bind_rows(Root_16S,Soil_16S) 
all_gyrB <- read_fasta("files_publication/AtSphere_gyrB.fasta") %>%
  as.data.frame() %>% 
  set_colnames("gyrB") %>% 
  rownames_to_column("Strain") 
all_rpoB <- read_fasta("files_publication/AtSphere_rpoB.fasta") %>%
  as.data.frame() %>% 
  set_colnames("rpoB") %>% 
  rownames_to_column("Strain") 
all_recA <- read_fasta("files_publication/AtSphere_recA.fasta") %>%
  as.data.frame() %>% 
  set_colnames("recA") %>% 
  rownames_to_column("Strain") 
all_dnaJ <- read_fasta("files_publication/AtSphere_dnaJ.fasta") %>%
  as.data.frame() %>% 
  set_colnames("dnaJ") %>% 
  rownames_to_column("Strain") 
all_atpD <- read_fasta("files_publication/AtSphere_aptD.fasta") %>%
  as.data.frame() %>% 
  set_colnames("atpD") %>% 
  rownames_to_column("Strain") 
all_thrC <- read_fasta("files_publication/AtSphere_thrC.fasta") %>%
  as.data.frame() %>% 
  set_colnames("thrC") %>% 
  rownames_to_column("Strain") 

phenotyped_set_concat <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
        filter(Strain %in% AUC_emmeans$Strain) %>% filter(Strain != "Root763")

relevant_concat_seqs <- phenotyped_set_concat %$% DNAStringSet(seq)
names(relevant_concat_seqs) <- phenotyped_set_concat$Strain



seqs_aligned <- AlignSeqs(relevant_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```


# Refined Taxonomic database

For reasons that are not completely clear, the order and class information is missing from the tables.
This information will be added from the bac120_taxonomy.tsv (genome taxonomy database https://www.nature.com/articles/nbt.4229) 

```{r eval = F}
tax_info <- read_tsv("files_publication/bac120_taxonomy_r95.tsv") %>%
  set_colnames(c("col1", "col2")) %>% 
  dplyr::select(col2) %>% 
  mutate(col2 = col2 %>% str_remove_all("[a-z]__")) %>% 
  separate(col2, c("Domain","Phylum","Class","Order","Family","Genus","Species"), sep = ";") %>% 
  dplyr::select(-Species) %>% 
  unique()
# Add missing information
tax_info <- rbind(tax_info,
            c("Bacteria",
              "Proteobacteria",
              "Betaproteobacteria",
              "Burkholderiales",
              "Comamonadaceae",
              "Acidovorax")) %>% 
            rbind(c("Bacteria",
               "Proteobacteria",
               "Gammaproteobacteria",
               "Xanthomonadales",
               "Xanthomonadaceae",
               "Pseudoxanthomonas")) %>% 
  # fix wrong taxonomy?
  mutate(Family = case_when(Genus == "Rhizobacter" ~ "Burkholderiales"),
         Class = case_when(Order == "Burkholderiales" ~ "Betaproteobacteria",
                           TRUE ~ Class))

write_rds(tax_info, "files_publication/taxonomy.rds")
```

# Read files

Read the growth data and the tree

```{r}
tax_info <- read_rds("files_publication/taxonomy.rds")



### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR <- read_rds("files_publication/Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
### read growth data
AUC_emmeans <- read_rds("files_publication/AUC_emmeans_pub.rds") %>%
  mutate(Family = case_when(Genus == "Rhizobacter" ~ "Burkholderiales",
                            TRUE ~ Family)) %>% 
  left_join(tax_info %>%
            dplyr::select(Genus, Class, Order), by = "Genus") %>% 
  filter(Strain %in% fit_GTR$tree$tip.label) ## Keep only those with sequences

tree_data <- AUC_emmeans %>% 
  mutate(id = str_c(Genus,Strain ,sep ="_"))

fit_GTR$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname

syncoms <-  rbind(
                  read_rds("files_publication/syncom_reduced_mixed.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Mixed"),
                 read_rds("files_publication/syncom_reduced_tolerant.rds") %>%
                   dplyr::select(Strain) %>% 
                   mutate(Syncom = "Reduced Tolerant"),
                 read_rds("files_publication/syncom_reduced_sensitive.rds") %>% 
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Reduced Sensitive"),
                 read_rds("files_publication/syncom_full_random.rds") %>%
                   dplyr::select(Strain) %>%
                   mutate(Syncom = "Random")
                 )
# Define colors for plotting
set.seed(42069)
family_colors <-  sample(c('#543005','#8c510a','#bf812d',
                           '#dfc27d','#f6e8c3','#4d9221',
                           '#c7eae5','#80cdc1','#35978f',
                           '#01665e','#003c30','#8e0152',
                           '#c51b7d','#de77ae','#f1b6da',
                           '#fde0ef','#000000','#e6f5d0',
                           '#b8e186','#7fbc41','#fb8072',
                           '#276419','#40004b','#bababa'),24)
names(family_colors) <- AUC_emmeans$Family %>% unique %>% sort
```

# Some descriptive stats of the growth data

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("1uM_APO","5uM_APO","10uM_APO","50uM_APO")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("1uM_APO","5uM_APO","10uM_APO","50uM_APO"))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  group_by(Treatment) %>% 
  mutate(Sensitive = case_when(estimate > 0.5 ~ "No",
                                   TRUE ~ "Yes")) %>% 
  group_by(Treatment, Sensitive) %>% 
  tally()
```

## Density plots

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("1uM_APO","5uM_APO","10uM_APO","50uM_APO")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("1uM_APO","5uM_APO","10uM_APO","50uM_APO"))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   #color = Treatment, 
                   fill = Treatment), alpha = 0.3) +
  theme_bw(base_family = "NimbusMon") +
  ggtitle("Estimated relative growth for APO treatments")
```

```{r}
AUC_emmeans %>% 
  filter(Treat.x %in% c("100uM_BOA","50uM_BOA", "10uM_BOA")) %>%
  dplyr::select(Strain, estimate, Treat.x) %>% 
  mutate(Treatment = fct_relevel(Treat.x, c("10uM_BOA","50uM_BOA","100uM_BOA" ))) %>% 
  dplyr::select(-Treat.x) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   #color = Treatment, 
                   fill = Treatment), alpha = 0.3) +
  theme_bw(base_family = "NimbusMon") +
    ggtitle("Estimated relative growth for BOA treatments")
```

```{r}
AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO") %>%
  dplyr::select(Strain, estimate, Phylum, Class, Order, Family, Genus) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   color = Class,
                   fill = Class), alpha = 0.4) +
  geom_density(aes(x=estimate), alpha = 1) +
  facet_wrap(~Phylum) +
  theme_bw(base_family = "NimbusMon") +
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "bottom") +
  ggtitle("Relative growth in 50uM APO") 
```
```{r}
AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO") %>%
  dplyr::select(Strain, estimate, Class, Order, Family, Genus) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   color = Family,
                   fill = Family), alpha = 0.4) +
  geom_density(aes(x=estimate, 
                   color = Family), alpha = 1) +
  geom_density(aes(x=estimate), alpha = 1) +
  facet_wrap(~fct_relevel(Class,
                          "Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli")) +
  scale_color_manual(values = family_colors) +
  scale_fill_manual(values = family_colors) +
  theme_bw(base_family = "Arial") +
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "none"
        ) +
  labs(title = "Relative growth in 50uM APO",
       x = "AUC[APO] / AUC[Control]") 
```
```{r}
AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO") %>%
  #dplyr::select(Strain, AUC_norm, Class, Order, Family, Genus) %>% 
  #distinct() %>% 
  # Drop Families with less than 2 members
  #filter(!Family %in% c("Moraxellaceae","Nocardiaceae","Mycobacteriaceae")) %>% 
  ggplot() +
  geom_point(aes(y = Family, 
                 x = estimate, 
                 fill = Family,
                 group = Strain),
             pch =21,
             alpha=0.3,
             size = 3,
             stat = "summary") +
  ggridges::geom_density_ridges(aes(x=AUC_norm, 
                                    y=Family,
                   color = Family,
                   fill = Family), 
                   alpha = 0.4) +
  facet_wrap(~fct_relevel(Class,
                          "Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli"),
             scales = "free_y") +
  scale_color_manual(values = family_colors) +
  scale_fill_manual(values = family_colors) +
  theme_bw(base_family = "Arial") +
  theme(strip.background = element_rect(fill = "white"), 
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "AUC[APO] / AUC[Control]") +
  lims(x=c(-0.3,1.6)) +
  NULL
```



```{r}

AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO", Phylum == "Proteobacteria") %>%
  dplyr::select(Strain, estimate, Class, Order, Family, Genus) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   color = Family,
                   fill = Family), alpha = 0.4) +
  geom_density(aes(x=estimate, 
                   color = Family), alpha = 0.4) +
  geom_density(aes(x=estimate), alpha = 1) +
    scale_color_manual(values = family_colors) +
  scale_fill_manual(values = family_colors) +
  facet_wrap(~Class) +
  theme_bw(base_family = "NimbusMon") +
  theme(strip.background = element_rect(fill = "white")) +
  ggtitle("Estimated relative growth for 50µM APO treatments")
```

```{r}
AUC_emmeans %>% 
  filter(Treat.x  == "50uM_APO") %>%
  dplyr::select(Strain, estimate, Class, Order, Family, Genus) %>% 
  distinct() %>% 
  ggplot() +
  geom_density(aes(x=estimate, 
                   #color = Treatment, 
                   fill = Family), alpha = 0.3) +
  facet_wrap(~Order) +
  theme_bw(base_family = "NimbusMon") +
  ggtitle("Estimated relative growth for APO treatments")
```

```{r}
AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50) %$% 
  var(AUC_norm)
```

```{r}
AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50) %>%
  group_by(Order) %>% 
  summarize(variance = var(AUC_norm))
AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50) %>%
  group_by(Family) %>% 
  summarize(variance = var(AUC_norm))
AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50) %>%
  group_by(Genus) %>% 
  summarize(variance = var(AUC_norm))
```

```{r}
AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50) %>%
  group_by(Family) %>% 
  summarize(variance = var(AUC_norm)) %$%
  range(variance)
  
```


```{r}
auc_apo_lm <- lm(AUC_norm ~ Class + Order + Family + Genus,
   data = AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50))
auc_apo_lm %>% anova
```

```{r}
summary(auc_apo_lm) 
```
```{r}
auc_apo_lm %>% emmeans::emmeans("Family")
```

Get variance, within, between

```{r}
library(nlme)
lme(AUC_norm~1, random=~1|Family,
   data = AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50)) %>% 
VarCorr()
# Within, between
lme(AUC_norm~1, random=~1|Class,
   data = AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50)) %>% 
VarCorr()
```
Residuals

```{r}
summary(auc_apo_lm)$sigma**2
```

```{r}
lm(AUC_norm ~ Family,
   data = AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50)) %>% 
  summary() %>% 
  vcov() %>% 
  diag()
```
```{r}
lm(AUC_norm ~ Genus,
   data = AUC_emmeans %>%
     filter(Chem == "APO", Conc == 50)) 
```


# Heatmap tree

```{r}
### Below creates "Group Info", which is currently the Genus. Extracted from tiplabels, see above chunk for modification of tiplabels.
tree_structure <- fit_GTR$tree %>% as.phylo
groupInfo <- split(fit_GTR$tree$tip.label, gsub("_\\w+", "", fit_GTR$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
```

```{r}
tree_plot_data <- cbind(
           tree_data %>%
             filter(Chem == "APO") %>% 
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id, Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             as.data.frame() %>% 
             dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
             separate(id, into = c("Genus","Strain"), sep = "_", remove = FALSE) %>% 
             left_join(., AUC_emmeans %>% dplyr::select(Strain, Family, Order, Class, Phylum), by = "Strain") %>% 
             distinct()  ,
          dummy1 = c(rep(NA,174)),
          dummy2 = c(rep(NA,174)),
         tree_data %>% filter(Chem == "BOA")  %>%
             mutate(Chem_conc = paste0(Chem,Conc)) %>% 
             dplyr::select(id,Chem_conc, estimate) %>%
             distinct %>% 
             spread(Chem_conc, estimate) %>%
             dplyr::select("BOA10","BOA50","BOA100") %>% 
             as.data.frame()) 



tree_data_boa <-  tree_plot_data %>% 
           dplyr::select("id","BOA10","BOA50","BOA100") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("10µM BOA", "50µM BOA", "100µM BOA"))
tree_data_apo <- tree_plot_data %>% 
           dplyr::select("id","APO1","APO5","APO10","APO50") %>% 
           column_to_rownames("id") %>% 
           set_colnames(c("1µM APO", "5µM APO", "10µM APO", "50µM APO"))
# Make a wide table of memberships, since syncoms were rbinded, there is one entry for each strain/syncom combination.


syncom_memberships <- left_join(tree_plot_data %>% dplyr::select(Strain, id),
  syncoms %>%
    mutate(member = as.factor(case_when(
                              Syncom == "Reduced Mixed" ~ "Mixed",
                              Syncom == "Reduced Tolerant" ~ "Tolerant",
                              Syncom == "Reduced Sensitive" ~ "Sensitive",
                              Syncom == "Random" ~ "Random",
                              TRUE ~ "NA"))) %>%
             pivot_wider(names_from = Syncom, values_from = member),
  by = "Strain")
```

### Family level heatmap

```{r}
fan_tree_family <- tree_structure %>%
  ggtree(branch.length = "none",
         layout = "fan",
         open.angle = 10) %<+% (tree_plot_data %>% mutate(Phylum_Family = paste(Phylum, Family, sep = "_")))

fan_tree_family <- fan_tree_family +
    geom_tippoint(aes(color =  Family,
                      shape = Class),
                  size = 1) +
  scale_color_manual("Family",
                     values = family_colors,
                     guide = "legend") +
  scale_shape(breaks = c("Alphaproteobacteria",
                         "Betaproteobacteria",
                         "Gammaproteobacteria",
                         "Actinomycetia",
                         "Bacteroidia",
                         "Bacilli"))
```

```{r}
fan_tree_family <- fan_tree_family + guides(Family = guide_legend(ncol = 2, byrow = TRUE))
```

### Add heatmaps

```{r}
heat_boa_fam <- gheatmap(fan_tree_family,
         tree_data_boa,
         offset=0.3,
         width=0.2,
         colnames_angle = 90,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth")  +
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth")
heat_boa_apo_fam <- gheatmap(heat_boa_fam, 
         tree_data_apo,
         offset=5.7,
         width=0.2,
         colnames_angle = 90,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon"))
heat_boa_apo_fam
```

### Full family plot

```{r}
fan_tree_family <- fan_tree_family + guides(Family = guide_legend(ncol = 3, byrow = TRUE))
heat_boa_fam <- gheatmap(fan_tree_family,
         tree_data_boa,
         offset=0.3,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth")  +
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth")
heat_boa_apo_fam <- gheatmap(heat_boa_fam, 
         tree_data_apo,
         offset=5.7,
         width=0.2,
         colnames_angle = 90,
         colnames = T, 
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon"))
heat_boa_apo_fam <- heat_boa_apo_fam + ggnewscale::new_scale_fill()
heat_boa_apo_fam_syncom <- gheatmap(heat_boa_apo_fam, 
         syncom_memberships %>% 
           dplyr::select("id","Random",
                              "Reduced Tolerant",
                              "Reduced Mixed",
                              "Reduced Sensitive") %>% 
           column_to_rownames("id"),
         offset=11.5,
         width=0.2,
         colnames_angle = 90,
         hjust = 0.5,
         font.size = 0) + 
  scale_fill_manual(na.value = "grey80",
                           name = "Syncom",
                           values = syncom_colors) +
  theme(text = element_text(family = "NimbusMon"),
        legend.position = "left") 
heat_boa_apo_fam_syncom + plot_annotation(title = "Fig 1") & theme(text = element_text(family = "NimbusMon")) 
```

### Plot without legend

```{r}
heat_boa_apo_fam_syncom +
    geom_tippoint(aes(color =  Family,
                      shape = Class),
                  size = 6) &
  theme(legend.position = "none") 
```

## Legend only

```{r}
cowplot::ggdraw(cowplot::get_legend(heat_boa_apo_fam_syncom))
```

# Family subset

```{r}
AUC_emmeans %$% Family %>% as.factor() %>% levels
```

```{r}
AUC_emmeans %>% dplyr::select(Family,Genus, Strain) %>% unique() %>% group_by(Family) %>% tally %>% arrange(desc(n), Family)
```

```{r}
AUC_emmeans %>% filter(Family == "Nocardioidaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Family == "Microbacteriaceae") %>% dplyr::select(Phylum) %>% unique()
```
```{r}
AUC_emmeans %>% filter(Phylum == "Firmicutes") %>% dplyr::select(Genus) %>% unique()
```

```{r}
AUC_emmeans %>% colnames
```

## Actinomycetia

```{r}
Actinomycetia <- AUC_emmeans %>% filter(Class == "Actinomycetia") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_actmy <- Actinomycetia$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Actinomycetia_tree <- drop.tip(tree_structure, not_actmy) 
```

```{r}
Actinomycetia_tree <- Actinomycetia_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
```


```{r}
Actinomycetia_tree_boa <- gheatmap(Actinomycetia_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish) +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Actinomycetia")+
  layout_dendrogram() +
    geom_tippoint(aes(color = Genus)) +
    ggthemes::scale_color_hc()
Actinomycetia_tree_boa_apo <- gheatmap(Actinomycetia_tree_boa, 
         tree_data_apo,
         offset=2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish)
```

```{r}
Actinomycetia_tree_boa_apo
```
## Commamodaceae

```{r}
Comamonadaceae <- AUC_emmeans %>% filter(Family == "Comamonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_comamo <- Comamonadaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Comamonadaceae_tree <- drop.tip(tree_structure, not_comamo) 
```

```{r}
Comamonadaceae_tree <- Comamonadaceae_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
```


```{r}
Comamonadaceae_tree_boa <- gheatmap(Comamonadaceae_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish) +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Comamonadaceae")+
  layout_dendrogram() +
    geom_tippoint(aes(color = Genus)) +
    ggthemes::scale_color_hc()
Comamonadaceae_tree_boa_apo <- gheatmap(Comamonadaceae_tree_boa, 
         tree_data_apo,
         offset=2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish)
```

```{r}
Comamonadaceae_tree_boa_apo
```


## Microbacteriaceae


```{r}
Microbacteria <- AUC_emmeans %>% filter(Family == "Microbacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Microbacteria <- Microbacteria$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Microbacteria_tree <- drop.tip(tree_structure, not_Microbacteria)

Microbacteria_tree <- Microbacteria_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Microbacteria_tree <- Microbacteria_tree + 
  layout_dendrogram() + 
  ggtitle("Microbacteria") +
    geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("Darjeeling2",4,"discrete"))

Microbacteria_tree <- rotate(Microbacteria_tree, 19)
```

```{r}
Microbacteria_tree_boa <- gheatmap(Microbacteria_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Microbacteriaceae, BOA") 


Microbacteria_tree_boa
```

```{r}
Microbacteria_tree_boa_apo <- gheatmap(Microbacteria_tree_boa, 
         tree_data_apo,
         offset=2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Microbacteriaceae") +
  theme(legend.position = "left")
Microbacteria_tree_boa_apo
```


## Rhizobiaceae


```{r}
Rhizobia <- AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_rhizobium <- Rhizobia$id %>% 
  {setdiff(tree_data$id %>% unique,.)}

```

```{r}
Rhizobia_tree <- drop.tip(tree_structure, not_rhizobium)
Rhizobia_tree <- Rhizobia_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Rhizobia_tree <- Rhizobia_tree + 
  layout_dendrogram() + 
  ggtitle("Rhizobia") +
    geom_tippoint(aes(color = Genus), size = 2) +

    scale_color_manual(values = wesanderson::wes_palette("BottleRocket2", 2 ,"discrete"))
Rhizobia_tree
```


```{r}
Rhizobia_tree_boa <- gheatmap(Rhizobia_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Rhizobiaceae, BOA") 


Rhizobia_tree_boa
```

```{r}
Rhizobia_tree_boa_apo <- gheatmap(Rhizobia_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Rhizobiaceae") +
  theme(legend.position = "left")
Rhizobia_tree_boa_apo
```


```{r}
Rhizobia_tree_apo <- gheatmap(Rhizobia_tree, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Rhizobiaceae, APO")
Rhizobia_tree_apo
```

```{r}
AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% filter(str_detect(Treat.x,"BOA")) %>% dplyr::select(Strain, Conc, estimate) %>% unique %>% filter(Conc == 100) %>% arrange(desc(estimate))
```

Root708 seems to grow particularily well in BOA

```{r}
AUC_emmeans %>% filter(Strain == "Root708") %>% 
  dplyr::select(Treat.x, estimate) %>% unique
```

## Xanthomonadaceae


```{r}
Xantho <- AUC_emmeans %>% filter(Family == "Xanthomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Xantho <- Xantho$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Xantho_tree <- drop.tip(tree_structure, not_Xantho)

Xantho_tree <- Xantho_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Xantho_tree <- Xantho_tree + 
  layout_dendrogram() + 
  ggtitle("Xanthomonadaceae") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Xantho_tree_boa <- gheatmap(Xantho_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Xanthoceae, BOA") 


Xantho_tree_boa
```

```{r}
Xantho_tree_boa_apo <- gheatmap(Xantho_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Xanthomonadaceae") +
  theme(legend.position = "left")
Xantho_tree_boa_apo
```


```{r}
Xantho_tree_apo <- gheatmap(Xantho_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Xanthomonadaceae, APO")
Xantho_tree_apo
```



## Bacillaceae


```{r}
Bacilli <- AUC_emmeans %>% 
  filter(Family == "Bacillaceae") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_Bacilli <- Bacilli$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Bacilli_tree <- drop.tip(tree_structure, not_Bacilli)

Bacilli_tree <- Bacilli_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Bacilli_tree <- Bacilli_tree + 
  layout_dendrogram() + 
  ggtitle("Bacillaceae") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Bacilli_tree_boa <- gheatmap(Bacilli_tree, 
         tree_data_boa,
         offset=0.16,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Bacilliceae, BOA") 


Bacilli_tree_boa
```

```{r}
Bacilli_tree_boa_apo <- gheatmap(Bacilli_tree_boa, 
         tree_data_apo,
         offset = 1.3,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillaceae") +
  theme(legend.position = "left")
Bacilli_tree_boa_apo
```


```{r}
Bacilli_tree_apo <- gheatmap(Bacilli_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillaceae, APO")
Bacilli_tree_apo
```

## Bacillales

All Firmicutes in the dataset are Bacillales (Paenibacillus, Bacillus)

```{r}
Firmi <- AUC_emmeans %>% 
  filter(Phylum == "Firmicutes") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_Firmi <- Firmi$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Firmi_tree <- drop.tip(tree_structure, not_Firmi)

Firmi_tree <- Firmi_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Firmi_tree <- Firmi_tree + 
  layout_dendrogram() + 
  ggtitle("Bacillales") +
  geom_tippoint(aes(color = Genus), size = 2) +
  scale_color_manual(values = wesanderson::wes_palette("IsleofDogs1", 3 ,"discrete"))
```


```{r}
Firmi_tree_boa <- gheatmap(Firmi_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Bacillales, BOA") 


Firmi_tree_boa
```

```{r}
Firmi_tree_boa_apo <- gheatmap(Firmi_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillales") +
  theme(legend.position = "left")
Firmi_tree_boa_apo
```


```{r}
Firmi_tree_apo <- gheatmap(Firmi_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillales, APO")
Firmi_tree_apo
```


## Nocardioidaceae


```{r}
Nocardio <- AUC_emmeans %>% filter(Family == "Nocardioidaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Nocardio <- Nocardio$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Nocardio_tree <- drop.tip(tree_structure, not_Nocardio)

Nocardio_tree <- Nocardio_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Nocardio_tree <- Nocardio_tree + 
  layout_dendrogram() + 
  ggtitle("Nocardio") +
    geom_tippoint(aes(color = Genus)) +
    scale_color_manual(values = wesanderson::wes_palette("GrandBudapest1", 2 ,"discrete"))
```


```{r}
Nocardio_tree_boa <- gheatmap(Nocardio_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Nocardioceae, BOA") 


Nocardio_tree_boa
```

```{r}
Nocardio_tree_boa_apo <- gheatmap(Nocardio_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Nocardioidaceae") +
  theme(legend.position = "left")
Nocardio_tree_boa_apo
```


```{r}
Nocardio_tree_apo <- gheatmap(Nocardio_tree, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Nocardioceae, APO")
Nocardio_tree_apo
```



## Pseudomonadaceae


```{r}
Pseudom <- AUC_emmeans %>% filter(Family == "Pseudomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_Pseudom <- Pseudom$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Pseudom_tree <- drop.tip(tree_structure, not_Pseudom)

Pseudom_tree <- Pseudom_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Pseudom_tree <- Pseudom_tree + 
  layout_dendrogram() + 
  ggtitle("Pseudom") +
    geom_tippoint(aes(color = Genus)) +
    scale_color_manual(values = wesanderson::wes_palette("Moonrise2", 2 ,"discrete"))
```


```{r}
Pseudom_tree_boa <- gheatmap(Pseudom_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames = FALSE,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Pseudomceae, BOA") 


Pseudom_tree_boa
```

```{r}
Pseudom_tree_boa_apo <- gheatmap(Pseudom_tree_boa, 
         tree_data_apo,
         offset=1.4,
         width=0.2,
         colnames = FALSE,
         colnames_angle = 0,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Pseudomonadaceae") +
  theme(legend.position = "left")
Pseudom_tree_boa_apo
```



## Alcaligenaceae

```{r}
Alcali <- AUC_emmeans %>% filter(Family == "Alcaligenaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_alcali <- Alcali$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Alcali_tree <- drop.tip(tree_structure, not_alcali)

Alcali_tree <- Alcali_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Alcali_tree <- Alcali_tree + 
  layout_dendrogram() + 
  ggtitle("Alcaligenaceae") +
    geom_tippoint(aes(color = Genus))
```


```{r}
Alcali_tree_boa <- gheatmap(Alcali_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Alcaligenaceae, BOA") 


Alcali_tree_boa
```

```{r}
Alcali_tree_boa_apo <- gheatmap(Alcali_tree_boa, 
         tree_data_apo,
         offset=0.7,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Alcaligenaceae") +
  theme(legend.position = "left")
Alcali_tree_boa_apo
```


## Flavobacteriaceae


```{r}
Flavobacteriaceae <- AUC_emmeans %>% filter(Family == "Flavobacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_flavo <- Flavobacteriaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Flavobacteriaceae_tree <- drop.tip(tree_structure, not_flavo)
```

```{r}
Flavobacteriaceae_tree <- Flavobacteriaceae_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Flavobacteriaceae_tree <- Flavobacteriaceae_tree + layout_dendrogram()
```


```{r}
Flavobacteriaceae_tree_boa <- gheatmap(Flavobacteriaceae_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Flavobacteriaceae") +
    geom_tippoint(aes(color = Genus)) +
    ggthemes::scale_color_hc()
Flavobacteriaceae_tree_boa_apo <- gheatmap(Flavobacteriaceae_tree_boa, 
         tree_data_apo,
         offset=0.8,
         width=0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth")
```
```{r}
Flavobacteriaceae_tree_boa_apo
```



```{r}
(Rhizobia_tree_boa_apo + Firmi_tree_boa_apo +
  plot_layout(guides = "collect")  &
  theme(legend.position = "left") &
  guides(fill =FALSE)) 
```

```{r}
((Rhizobia_tree_boa_apo + Firmi_tree_boa_apo )/
   (Pseudom_tree_boa_apo + Alcali_tree_boa_apo)) &
  theme(legend.position = "bottom") &
  guides(fill =FALSE)
```

# Order Trees

## Rhizobiales

```{r}
Rhizobia <- AUC_emmeans %>% filter(Order == "Rhizobiales") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
```

```{r}
not_rhizobia <- Rhizobia$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Rhizobia_tree <- drop.tip(tree_structure, not_rhizobia)
```

```{r}
Rhizobia_tree <- Rhizobia_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Rhizobia_tree <- Rhizobia_tree + 
  layout_dendrogram() + 
  ggtitle("Rhizobiales") +
  geom_tippoint(aes(color = Genus), size = 2)
```


```{r}
Rhizobia_tree_boa <- gheatmap(Rhizobia_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Rhizobiales, BOA") 


Rhizobia_tree_boa
```

```{r}
Rhizobia_tree_boa_apo <- gheatmap(Rhizobia_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Rhizobiales") +
  theme(legend.position = "left")
Rhizobia_tree_boa_apo
```

## Xanthomodales

```{r}
Xanthos <- AUC_emmeans %>%
  filter(Order == "Xanthomonadales") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_xanthos <- Xanthos$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Xanthos_tree <- drop.tip(tree_structure, not_xanthos)
```

```{r}
Xanthos_tree <- Xanthos_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Xanthos_tree <- Xanthos_tree + 
  layout_dendrogram() + 
  ggtitle("Xanthomodales") +
  geom_tippoint(aes(color = Genus), size = 2)
```


```{r}
Xanthos_tree_boa <- gheatmap(Xanthos_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) +
  ggtitle("Xanthomodales, BOA") 


Xanthos_tree_boa
```

```{r}
Xanthos_tree_boa_apo <- gheatmap(Xanthos_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Xanthomodales") +
  theme(legend.position = "left")
Xanthos_tree_boa_apo
```
## Bacillales

```{r}
Bacillales <- AUC_emmeans %>%
  filter(Order == "Bacillales") %>%
  dplyr::select(Genus, Strain) %>%
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
```

```{r}
not_bacilla <- Bacillales$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
Bacillales_tree <- drop.tip(tree_structure, not_bacilla)
```

```{r}
Bacillales_tree <- Bacillales_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
Bacillales_tree <- Bacillales_tree + 
  layout_dendrogram() + 
  ggtitle("Xanthomodales") +
  geom_tippoint(aes(color = Genus), size = 2)
```


```{r}
Bacillales_tree_boa <- gheatmap(Bacillales_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

```

```{r}
Bacillales_tree_boa_apo <- gheatmap(Bacillales_tree_boa, 
         tree_data_apo,
         offset=1.8,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         colnames_position = "bottom",
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacillales") +
  theme(legend.position = "left")
Bacillales_tree
```


# Phylosig

```{r}
tree_data_physig <- tree_data %>%
             filter(Chem == "APO", Conc == 50) %>% 
             dplyr::select(id, estimate, std.error) %>%
             distinct %>% 
             as.data.frame() %>% 
             separate("id", c("Genus","Strain"), sep = "_", remove = F) %>% 
             column_to_rownames("id")
all_se <- tree_data_physig$std.error
names(all_se) <- rownames(tree_data_physig)
```


## All




```{r}
apo_50um_sorted <- motmot::sortTraitData(phy = fit_GTR$tree %>%
                                   multi2di() ,
                                 y = tree_data_physig,
                                 data.name = "estimate",
                                 log.trait = FALSE,
                                 pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted$phy,
            x = apo_50um_sorted$trait,
            method = "K",
            se = all_se,
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted$phy,
            x = apo_50um_sorted$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)

```
## Gram pos

```{r}
grampos_strains <- AUC_emmeans %>%
  filter(Class %in% c("Actinomycetia","Bacilli")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_grampos <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
  filter(Strain %in% grampos_strains) %>%
  filter(Strain != "Root763") 
  

grampos_concat_seqs <- phenotyped_set_concat_grampos %$% DNAStringSet(seq)
names(grampos_concat_seqs) <- phenotyped_set_concat_grampos$Strain
```

```{r eval =F}
seqs_aligned <- AlignSeqs(grampos_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Grampos_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
tree_data_grampos <- tree_data_physig %>% filter(Strain %in% grampos_strains)

### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_grampos <- read_rds("files_publication/Grampos_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_grampos$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_grampos <- motmot::sortTraitData(phy = fit_GTR_grampos$tree %>%
                                                     multi2di() ,
                                                   y = tree_data_physig,
                                                   data.name = "estimate",
                                                   log.trait = FALSE,
                                                   pass.ultrametric = TRUE)
```


```{r}
phytools::phylosig(tree = apo_50um_sorted_grampos$phy,
            x = apo_50um_sorted_grampos$trait,
            method = "K",
            se = all_se,
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_grampos$phy,
            x = apo_50um_sorted_grampos$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```
### Tree

```{r}
tree_structure <- fit_GTR_grampos$tree %>% as.phylo
groupInfo <- split(fit_GTR_grampos$tree$tip.label, gsub("_\\w+", "", fit_GTR_grampos$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Actinomycetia", "Bacilli" )) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(fit_GTR_grampos$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family, shape = Class), size = 4) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Gram positives") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Gram positives, APO")
grampos_tree <- subset_tree_boa_apo
```

```{r}
grampos_tree +
  theme(text = element_text(family = "NimbusMon"))
```


## Actinomycetia

```{r}
actinomycetia_strains <- AUC_emmeans %>%
  filter(Class %in% c("Actinomycetia")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_actinomyc <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
  filter(Strain %in% actinomycetia_strains) %>%
  filter(Strain != "Root763") 
  

actinomyc_concat_seqs <- phenotyped_set_concat_actinomyc %$% DNAStringSet(seq)
names(actinomyc_concat_seqs) <- phenotyped_set_concat_actinomyc$Strain
```

```{r eval =F}
seqs_aligned <- AlignSeqs(actinomyc_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Actinomycetia_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_actinomyc <- read_rds("files_publication/Actinomycetia_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_actinomyc$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_actinomyc <- motmot::sortTraitData(phy = fit_GTR_actinomyc$tree %>%
                                                     multi2di() ,
                                                   y = tree_data_physig,
                                                   data.name = "estimate",
                                                   log.trait = FALSE,
                                                   pass.ultrametric = TRUE)
```


```{r}
phytools::phylosig(tree = apo_50um_sorted_actinomyc$phy,
            x = apo_50um_sorted_actinomyc$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

#### Actinomycetales

```{r}
Actinomycetales <- AUC_emmeans %>% filter(Order == "Actinomycetales") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
tree_data_actinomycetales <- tree_data_physig %>% filter(Strain %in% Actinomycetales$Strain)

apo_50um_sorted_actinomycetales <- motmot::sortTraitData(phy = fit_GTR_actinomyc$tree %>% multi2di() ,
                                              y = tree_data_actinomycetales,
                                              data.name = "estimate", 
                                              log.trait = FALSE,
                                              pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_actinomycetales$phy,
            x = apo_50um_sorted_actinomycetales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

#### Intrasporangiaceae

```{r}
Intraspo <- AUC_emmeans %>% filter(Family == "Intrasporangiaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
tree_data_intraspo <- tree_data_physig %>% filter(Strain %in% Intraspo$Strain)

apo_50um_sorted_intraspo <- motmot::sortTraitData(phy = fit_GTR_actinomyc$tree %>% multi2di() ,
                                              y = tree_data_intraspo,
                                              data.name = "estimate", 
                                              log.trait = FALSE,
                                              pass.ultrametric = TRUE)

```


```{r}
phytools::phylosig(tree = apo_50um_sorted_intraspo$phy,
            x = apo_50um_sorted_intraspo$trait,
            method = "lambda",
            #se = all_se,
            test = TRUE)
```
#### Nocardioidaceae

```{r}
Nocardioid <- AUC_emmeans %>% filter(Family == "Nocardioidaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
tree_data_nocardioid <- tree_data_physig %>% filter(Strain %in% Nocardioid$Strain)

apo_50um_sorted_nocardioid <- motmot::sortTraitData(phy = fit_GTR_actinomyc$tree %>% multi2di() ,
                                              y = tree_data_nocardioid,
                                              data.name = "estimate", 
                                              log.trait = FALSE,
                                              pass.ultrametric = TRUE)

```


```{r}
phytools::phylosig(tree = apo_50um_sorted_nocardioid$phy,
            x = apo_50um_sorted_nocardioid$trait,
            method = "lambda",
            #se = all_se,
            test = TRUE)
```
#### Microbacteriaceae

```{r}
Microbacteria <- AUC_emmeans %>% filter(Family == "Microbacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
not_Microbacteria <- Microbacteria$id %>% 
  {setdiff(tree_data$id %>% unique,.)}

apo_50um_sorted_microbacteriacae <- motmot::sortTraitData(phy = fit_GTR_actinomyc$tree %>% 
                                        drop.tip(not_Microbacteria) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_microbacteriacae$phy,
            x = apo_50um_sorted_microbacteriacae$trait,
            method = "lambda",
            #se = all_se,
            test = TRUE)
```


### Tree

```{r}
tree_structure <- fit_GTR_actinomyc$tree %>% as.phylo
groupInfo <- split(fit_GTR_actinomyc$tree$tip.label, gsub("_\\w+", "", fit_GTR_actinomyc$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Actinomycetia" )) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data_actinomyc_apo$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family, shape = Class), size = 4) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Actinomycetia, APO")
actinomyc_tree <- subset_tree_boa_apo
```

```{r}
actinomyc_tree
```

## Bacilli

```{r}
baci_strains <- AUC_emmeans %>%
  filter(Class %in% c("Bacilli")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_baci <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
  filter(Strain %in% baci_strains) %>%
  filter(Strain != "Root763") 
  

baci_concat_seqs <- phenotyped_set_concat_baci %$% DNAStringSet(seq)
names(baci_concat_seqs) <- phenotyped_set_concat_baci$Strain
```

```{r eval =F}
seqs_aligned <- AlignSeqs(baci_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Bacilli_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
## read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_baci <- read_rds("files_publication/Bacilli_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_baci$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_baci <- motmot::sortTraitData(phy = fit_GTR_baci$tree %>%
                                                     multi2di() ,
                                                   y = tree_data_physig,
                                                   data.name = "estimate",
                                                   log.trait = FALSE,
                                                   pass.ultrametric = TRUE)
```


```{r}
phytools::phylosig(tree = apo_50um_sorted_baci$phy,
            x = apo_50um_sorted_baci$trait,
            se = all_se,
            method = "K",
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_baci$phy,
            x = apo_50um_sorted_baci$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```
#### Bacillales

```{r}
apo_50um_sorted_bacila <- motmot::sortTraitData(phy = fit_GTR_grampos$tree %>% 
                                        drop.tip(not_bacilla) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_bacila$phy,
            x = apo_50um_sorted_bacila$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```

### Tree

```{r}
tree_structure <- fit_GTR_baci$tree %>% as.phylo
groupInfo <- split(fit_GTR_baci$tree$tip.label, gsub("_\\w+", "", fit_GTR_baci$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Bacilli" )) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data_baci_apo$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family, shape = Class), size = 4) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Bacilli, APO")
bacilli_tree <- subset_tree_boa_apo
```

```{r}
bacilli_tree
```

## Proteobacteria

```{r}
proteobacteria_strains <- AUC_emmeans %>%
  filter(Class %in% c("Alphaproteobacteria", "Betaproteobacteria", "Gammaproteobacteria" )) %$%
  Strain %>% 
  unique
phenotyped_set_concat_proteo <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
  filter(Strain %in% proteobacteria_strains) %>%
  filter(Strain != "Root763") 
  

proteo_concat_seqs <- phenotyped_set_concat_proteo %$% DNAStringSet(seq)
names(proteo_concat_seqs) <- phenotyped_set_concat_proteo$Strain
```


```{r eval =F}
seqs_aligned <- AlignSeqs(proteo_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_proteo <- read_rds("files_publication/Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_proteo$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```


```{r}
apo_50um_sorted_proteo <- motmot::sortTraitData(phy = fit_GTR_proteo$tree %>%
                                   multi2di() ,
                                 y = tree_data_physig,
                                 data.name = "estimate",
                                 log.trait = FALSE,
                                 pass.ultrametric = TRUE)
```


```{r}
phytools::phylosig(tree = apo_50um_sorted_proteo$phy,
            x = apo_50um_sorted_proteo$trait,
            se = all_se,
            method = "K",
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_proteo$phy,
            x = apo_50um_sorted_proteo$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```

### Tree

```{r}
tree_structure <- fit_GTR_proteo$tree %>% as.phylo
groupInfo <- split(fit_GTR_proteo$tree$tip.label, gsub("_\\w+", "", fit_GTR_proteo$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Alphaproteobacteria", "Betaproteobacteria", "Gammaproteobacteria" )) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data_proteo_apo$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family, shape = Class), size = 4) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Proteobacteria, APO")
proteo_tree <- subset_tree_boa_apo
```

```{r}
proteo_tree
```

```{r}
caulo <- AUC_emmeans %>% filter(Family == "Caulobacteraceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
not_caulo <- caulo$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
caulo_tree <- drop.tip(tree_structure, not_caulo)

caulo_tree <- caulo_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
caulo_tree <- caulo_tree + 
  layout_dendrogram() + 
  geom_tiplab(angle = 30) 
caulo_tree
```
```{r}
xanth <- AUC_emmeans %>% filter(Family == "Xanthomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
not_xanth <- xanth$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
xanth_tree <- drop.tip(tree_structure, not_xanth)

xanth_tree <- xanth_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
xanth_tree <- xanth_tree + 
  layout_dendrogram() + 
  geom_tiplab(angle = 30) 
xanth_tree
```
```{r}
sphing <- AUC_emmeans %>% filter(Family == "Sphingomonadaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
not_sphing <- sphing$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
sphing_tree <- drop.tip(tree_structure, not_sphing)

sphing_tree <- sphing_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
sphing_tree <- sphing_tree + 
  layout_dendrogram() + 
  geom_tiplab(angle = 30) 
sphing_tree
```

## Betaproteo

```{r}
## Proteobacteria

bproteobacteria_strains <- AUC_emmeans %>%
  filter(Class %in% c("Betaproteobacteria")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_bproteo <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
  filter(Strain %in% bproteobacteria_strains) %>% filter(Strain != "Root763") 
  

bproteo_concat_seqs <- phenotyped_set_concat_bproteo %$% DNAStringSet(seq)
names(bproteo_concat_seqs) <- phenotyped_set_concat_bproteo$Strain
```


```{r eval =F}
seqs_aligned <- AlignSeqs(bproteo_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Beta_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")

seqs_aligned <- AlignSeqs(gproteo_concat_seqs, anchor=NA)
```

```{r}
### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_bproteo <- read_rds("files_publication/Beta_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_bproteo$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_beta <- motmot::sortTraitData(phy = fit_GTR_bproteo$tree %>% multi2di() ,
                                              y = tree_data_physig,
                                              data.name = "estimate", 
                                              log.trait = FALSE,
                                              pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_beta$phy,
            x = apo_50um_sorted_beta$trait,
            se = all_se,
            method = "K",
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_beta$phy,
            x = apo_50um_sorted_beta$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```
#### Comamonadaceae

```{r}
apo_50um_sorted_comamo <- motmot::sortTraitData(phy = fit_GTR_bproteo$tree %>% 
                                                  drop.tip(not_comamo)  %>% 
                                                  multi2di() ,
                                              y = tree_data_physig,
                                              data.name = "estimate", 
                                              log.trait = FALSE,
                                              pass.ultrametric = TRUE)

```

```{r}
phytools::phylosig(tree = apo_50um_sorted_comamo$phy,
            x = apo_50um_sorted_comamo$trait,
            method = "K",
            se = all_se,
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_comamo$phy,
            x = apo_50um_sorted_comamo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```
```{r}
tree_data_comamo %>% plot_tree
```



#### Tree

```{r}
tree_structure <- fit_GTR_bproteo$tree %>% as.phylo
groupInfo <- split(fit_GTR_bproteo$tree$tip.label, gsub("_\\w+", "", fit_GTR_bproteo$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Betaproteobacteria")) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family), size = 2) 


subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Beta Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Beta Proteobacteria, APO")
bproteo_tree <- subset_tree_boa_apo
```

```{r}
bproteo_tree
```

### Burkholderiales

```{r}
apo_50uM_burkholderiales <- motmot::sortTraitData(phy = fit_GTR_bproteo$tree %>% multi2di() ,
                                               y = tree_data_physig %>% 
                        filter(Strain %in% (AUC_emmeans %>%
                                              filter(Order == "Burkholderiales") %$% 
                                              Strain %>%
                                              unique)), 
                                               data.name = "estimate",
                                               log.trait = F)
```

```{r}
phytools::phylosig(tree = apo_50uM_burkholderiales$phy,
            x = apo_50uM_burkholderiales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

## Alphaproteo

```{r}
## Alphaproteo

aproteobacteria_strains <- AUC_emmeans %>%
  filter(Class %in% c("Alphaproteobacteria")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_aproteo <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>%   
  filter(Strain %in% aproteobacteria_strains) %>% filter(Strain != "Root763") 
  

aproteo_concat_seqs <- phenotyped_set_concat_aproteo %$% DNAStringSet(seq)
names(aproteo_concat_seqs) <- phenotyped_set_concat_aproteo$Strain
```


```{r eval =F}
seqs_aligned <- AlignSeqs(aproteo_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Alpha_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_aproteo <- read_rds("files_publication/Alpha_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_aproteo$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_alpha <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                         multi2di(),
                                         y = tree_data_physig, 
                                         data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_alpha$phy,
            x = apo_50um_sorted_alpha$trait,
            method = "K",
            se = all_se,
            test = TRUE)
```
```{r}
phytools::phylosig(tree = apo_50um_sorted_alpha$phy,
            x = apo_50um_sorted_alpha$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```

### Tree

```{r}
tree_structure <- fit_GTR_aproteo$tree %>% as.phylo
groupInfo <- split(fit_GTR_aproteo$tree$tip.label, gsub("_\\w+", "", fit_GTR_aproteo$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Alphaproteobacteria")) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop) 

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family), size = 2) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Alpha Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Alpha Proteobacteria, APO")
aproteo_tree <- subset_tree_boa_apo
```

```{r}
aproteo_tree
```
### Rhizobiales

```{r}
Rhizobia <- AUC_emmeans %>% filter(Order == "Rhizobiales") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique

Rhizobiaceae <- AUC_emmeans %>% filter(Family == "Rhizobiaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
Rhizobium <- AUC_emmeans %>% filter(Genus == "Rhizobium") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique
Ensifer <- AUC_emmeans %>% filter(Genus == "Ensifer") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique


not_rhizobia <- Rhizobia$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
not_rhizobiaceae <- Rhizobiaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
not_rhizobium <- Rhizobium$id %>% 
  {setdiff(tree_data$id %>% unique,.)}
not_ensifer <- Ensifer$id %>% 
  {setdiff(tree_data$id %>% unique,.)}

apo_50um_sorted_rhizobiales <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                        drop.tip(not_rhizobia) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
apo_50um_sorted_rhizobium <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                        drop.tip(not_rhizobium) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
apo_50um_sorted_rhizobiaceae <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                        drop.tip(not_rhizobiaceae) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
apo_50um_sorted_ensifer <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                        drop.tip(not_rhizobium) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
```


```{r}
phytools::phylosig(tree = apo_50um_sorted_rhizobiales$phy,
            x = apo_50um_sorted_rhizobiales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```
### Rhizobiaceae

```{r}
phytools::phylosig(tree = apo_50um_sorted_rhizobiaceae$phy,
            x = apo_50um_sorted_rhizobiaceae$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```
#### Rhizobium

```{r}
phytools::phylosig(tree = apo_50um_sorted_rhizobium$phy,
            x = apo_50um_sorted_rhizobium$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

#### Ensifer

```{r}
phytools::phylosig(tree = apo_50um_sorted_ensifer$phy,
            x = apo_50um_sorted_ensifer$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

### Phyllobacteriaceae


```{r}
Phyllobacteriaceae <- AUC_emmeans %>% filter(Family == "Phyllobacteriaceae") %>% dplyr::select(Genus, Strain) %>% mutate(id = paste(Genus, Strain, sep = "_")) %>% unique



not_phyllo <- Phyllobacteriaceae$id %>% 
  {setdiff(tree_data$id %>% unique,.)}

apo_50um_sorted_phyllo <- motmot::sortTraitData(phy = fit_GTR_aproteo$tree %>% 
                                        drop.tip(not_phyllo) %>%
                                         multi2di(),
                                        y = tree_data_physig,
                                        data.name = "estimate",
                                       log.trait = FALSE,
                                       pass.ultrametric = TRUE)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_phyllo$phy,
            x = apo_50um_sorted_phyllo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
```

## Gammaproteo

```{r}
## Proteobacteria

gproteobacteria_strains <- AUC_emmeans %>%
  filter(Class %in% c("Gammaproteobacteria")) %$%
  Strain %>% 
  unique
phenotyped_set_concat_gproteo <- left_join(all_16s, all_gyrB, by = "Strain") %>% 
  left_join(all_rpoB, by = "Strain") %>% 
  left_join(all_recA, by = "Strain") %>% 
  left_join(all_dnaJ, by = "Strain") %>% 
  left_join(all_atpD, by = "Strain") %>% 
  left_join(all_thrC, by = "Strain") %>% 
  mutate(seq = paste0(rRNA16S,gyrB,rpoB,recA,dnaJ,atpD,thrC)) %>% 
   filter(Strain %in% gproteobacteria_strains) %>% filter(Strain != "Root763") 
  

gproteo_concat_seqs <- phenotyped_set_concat_gproteo %$% DNAStringSet(seq)
names(gproteo_concat_seqs) <- phenotyped_set_concat_gproteo$Strain
```


```{r eval =F}
seqs_aligned <- AlignSeqs(gproteo_concat_seqs, anchor=NA)

### Make tree
phang_align <- phyDat(as(seqs_aligned, "matrix"), type="DNA")
dist_m <- dist.ml(phang_align)
tree_NJ <- NJ(dist_m) #  Note, tip order != sequence order
fit = pml(tree_NJ, data=phang_align)
### Make GTR tree
fit_GTR <- update(fit, k=4, inv=0.2)
fit_GTR <- optim.pml(fit_GTR, model="GTR", optInv=TRUE, optGamma=TRUE,
 rearrangement = "stochastic", control = pml.control(trace = 0))
write_rds(fit_GTR, "files_publication/Gamma_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
```

```{r}
### read tree
#fit_GTR <- read_rds("files_publication/Strain_tree.rds")
fit_GTR_gproteo <- read_rds("files_publication/Gamma_Proteobacteria_Strain_tree_16S_gyrB_rhoB_recA_dnaJ_atpD_thrC.rds")
fit_GTR_gproteo$tree$tip.label %<>%
  as.data.frame() %>% 
  set_colnames("Strain") %>%
  left_join(., ( AUC_emmeans %>% dplyr::select(Strain, Genus) %>% distinct)) %>%
  mutate(Strain = str_c(Genus,Strain, sep ="_")) %>%
  dplyr::select(-Genus) %>%
  as.vector() %>%
  unlist %>%
  unname
```

```{r}
apo_50um_sorted_gamma <- motmot::sortTraitData(phy = fit_GTR_gproteo$tree %>% multi2di() ,
                                               y = tree_data_physig, 
                                               data.name = "estimate",
                                               log.trait = F)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_gamma$phy,
            x = apo_50um_sorted_gamma$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```
#### Xanthomonadales

```{r}
apo_50um_sorted_xantho <- motmot::sortTraitData(phy = fit_GTR_gproteo$tree %>%
                                                  drop.tip(not_xanthos) %>%
                                                  multi2di() ,
                                               y = tree_data_physig, 
                                               data.name = "estimate",
                                               log.trait = F)
```

```{r}
phytools::phylosig(tree = apo_50um_sorted_xantho$phy,
            x = apo_50um_sorted_xantho$trait,
            se = all_se,
            method = "lambda",
            test = TRUE) 
```
```{r}
Xanthos_tree_boa_apo +
  ggtitle("Xanthomonadales",
          sub = expression(paste(lambda, " = 1, p < 0.001" )))
```

### Tree

```{r}
tree_structure <- fit_GTR_gproteo$tree %>% as.phylo
groupInfo <- split(fit_GTR_gproteo$tree$tip.label, gsub("_\\w+", "", fit_GTR_gproteo$tree$tip.label))
tree_structure <- groupOTU(tree_structure, groupInfo)
ids_to_keep <- AUC_emmeans %>%
  filter(Class %in% c("Gammaproteobacteria" )) %>%
  dplyr::select(Genus, Strain) %>% 
  mutate(id = paste(Genus, Strain, sep = "_")) %>%
  unique
drop <-  setdiff(tree_data$id %>% unique,ids_to_keep$id)
subset_tree <- drop.tip(tree_structure, drop)

subset_tree <- subset_tree %>%
  ggtree(branch.length = "none",
         layout = "rectangular",
         open.angle = 10) %<+% tree_plot_data
subset_tree <- subset_tree + 
  layout_dendrogram() + 
  geom_tippoint(aes(color = Family), size = 2) 

subset_tree_boa <- gheatmap(subset_tree, 
         tree_data_boa,
         offset=0.2,
         width=0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  theme(text = element_text(family = "NimbusMon")) 

subset_tree_boa_apo <- gheatmap(subset_tree_boa, 
         tree_data_apo,
         offset = 1.55,
         width = 0.2,
         colnames_angle = 0,
         colnames = FALSE,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Gamma Proteobacteria") +
  theme(legend.position = "left")

subset_tree_boa_apo <- gheatmap(subset_tree, 
         tree_data_apo,
         offset = 0.8,
         width = 0.2,
         hjust = 1,
         font.size = 3,
         legend_title = "Rel. growth") + 
  scale_fill_viridis_c(na.value = "white",
                       direction = -1,
                       limits = c(0, 1.5),
                       oob = scales::squish,
                       name = "Rel. growth") +
  ggtitle("Gamma Proteobacteria, APO")
gproteo_tree <- subset_tree_boa_apo
```

```{r}
gproteo_tree
```



# Only Tests

## All

```{r}
lambda_all <- phytools::phylosig(tree = apo_50um_sorted$phy,
            x = apo_50um_sorted$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_all
```

## Proteo

```{r}
lambda_proteo <- phytools::phylosig(tree = apo_50um_sorted_proteo$phy,
            x = apo_50um_sorted_proteo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_proteo
```

### Alpha

```{r}
lambda_alpha <- phytools::phylosig(tree = apo_50um_sorted_alpha$phy,
            x = apo_50um_sorted_alpha$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
lambda_alpha
```
#### Rhizobiales

```{r}
lambda_rhizo <- phytools::phylosig(tree = apo_50um_sorted_rhizobiales$phy,
            x = apo_50um_sorted_rhizobiales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_rhizo
```

##### Rhizobiaceae

```{r}
lambda_rhizobiaceae <- phytools::phylosig(tree = apo_50um_sorted_rhizobiaceae$phy,
            x = apo_50um_sorted_rhizobiaceae$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_rhizobiaceae
```
##### Phyllobacteriaceae

```{r}
lambda_phyllo <- phytools::phylosig(tree = apo_50um_sorted_phyllo$phy,
            x = apo_50um_sorted_phyllo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_phyllo
```



### beta

```{r}
lambda_beta <- phytools::phylosig(tree = apo_50um_sorted_beta$phy,
            x = apo_50um_sorted_beta$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
lambda_beta
```
#### Comamonadaceae

```{r}
lambda_comamo <- phytools::phylosig(tree = apo_50um_sorted_comamo$phy,
            x = apo_50um_sorted_comamo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_comamo
```

#### Burkholderiales

```{r}
lambda_burk <- phytools::phylosig(tree = apo_50uM_burkholderiales$phy,
            x = apo_50uM_burkholderiales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_burk
```
### gamma

```{r}
lambda_gamma <- phytools::phylosig(tree = apo_50um_sorted_gamma$phy,
            x = apo_50um_sorted_gamma$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
lambda_gamma
```
#### Xanthomonadales

```{r}
lambda_xantho <- phytools::phylosig(tree = apo_50um_sorted_xantho$phy,
            x = apo_50um_sorted_xantho$trait,
            se = all_se,
            method = "lambda",
            test = TRUE) 
lambda_xantho
```
```{r}
lambda_xanthomonadaceae <- phytools::phylosig(tree = apo_50um_sorted_xantho$phy,
            x = apo_50um_sorted_xantho$trait,
            se = all_se,
            method = "lambda",
            test = TRUE) 
lambda_xantho
```



## Gram pos

```{r}
lambda_gpos <- phytools::phylosig(tree = apo_50um_sorted_grampos$phy,
            x = apo_50um_sorted_grampos$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_gpos
```

### Actinomycetia

```{r}
lambda_actinomy <- phytools::phylosig(tree = apo_50um_sorted_actinomyc$phy,
            x = apo_50um_sorted_actinomyc$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_actinomy
```
#### Actinomycetales 

```{r}
lambda_actinomycetales <- phytools::phylosig(tree = apo_50um_sorted_actinomycetales$phy,
            x = apo_50um_sorted_actinomycetales$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_actinomycetales
```


#### Intrasporangiaceae

```{r}
lambda_intraspo <- phytools::phylosig(tree = apo_50um_sorted_intraspo$phy,
            x = apo_50um_sorted_intraspo$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_intraspo
```


#### Nocardioidaceae

```{r}
lambda_nocardioid <- phytools::phylosig(tree = apo_50um_sorted_nocardioid$phy,
            x = apo_50um_sorted_nocardioid$trait,
            method = "lambda",
            #se = all_se,
            test = TRUE)
lambda_nocardioid
```

#### Microbacteriaceae

```{r}
lambda_microbact <- phytools::phylosig(tree = apo_50um_sorted_microbacteriacae$phy,
            x = apo_50um_sorted_microbacteriacae$trait,
            method = "lambda",
            se = all_se,
            test = TRUE)
lambda_microbact
```


### Bacilli

```{r}
lambda_baci <- phytools::phylosig(tree = apo_50um_sorted_baci$phy,
            x = apo_50um_sorted_baci$trait,
            se = all_se,
            method = "lambda",
            test = TRUE)
```

### Tab S2

```{r}
lambda_tab <- rbind(c(lambda_all$lambda %>% round(3), lambda_all$P %>% round(4),
                      "All", "All", "All", "All","All"),
                    c(lambda_gpos$lambda %>% round(3), lambda_gpos$P %>% round(4),
                      "Gram Positive","All" ,"All", "All", "All"),
                    c(lambda_actinomy$lambda %>% round(3), lambda_actinomy$P %>% round(4),
                      "Gram Positive", "Actinobacteria", "Actinomycetia", "All", "All"),
                    c(lambda_actinomycetales$lambda %>% round(3), lambda_actinomycetales$P %>% round(4), 
                      "Gram Positive", "Actinobacteria", "Actinomycetia", "Actinomycetales", "All"),
                    c(lambda_nocardioid$lambda %>% round(3), lambda_nocardioid$P %>% round(4), 
                      "Gram Positive", "Actinobacteria", "Actinomycetia", "Actinomycetales", "Nocardioidaceae"),
                    c(lambda_intraspo$lambda %>% round(3), lambda_intraspo$P %>% round(4), 
                      "Gram Positive", "Actinobacteria", "Actinomycetia", "Actinomycetales", "Intrasporangiaceae"),
                    c(lambda_microbact$lambda %>% round(3), lambda_microbact$P %>% round(4), 
                      "Gram Positive", "Actinobacteria", "Actinomycetia", "Actinomycetales", "Microbacteriaceae"),
                    c(lambda_baci$lambda %>% round(3), lambda_baci$P %>% round(4), 
                      "Gram Positive", "Firmicutes", "Bacilli", "All", "All"),
                    c(lambda_proteo$lambda %>% round(3), lambda_proteo$P %>% round(4),
                      "Gram Negative", "Proteobacteria", "All", "All", "All"),
                    c(lambda_alpha$lambda %>% round(3), lambda_alpha$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Alpha", "All", "All"),
                    c(lambda_rhizo$lambda %>% round(3), lambda_rhizo$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Alpha", "Rhizobiales", "All"),
                    c(lambda_rhizobiaceae$lambda %>% round(3), lambda_rhizobiaceae$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Alpha", "Rhizobiales", "Rhizobiacaea"),
                    c(lambda_phyllo$lambda %>% round(3), lambda_phyllo$P %>% round(4),
                      "Gram Negative", "Proteobacteria", "Alpha", "Rhizobiales", "Phyllobacteriacae"),
                    c(lambda_beta$lambda %>% round(3), lambda_beta$P %>% round(4),
                      "Gram Negative", "Proteobacteria", "Beta", "All", "All"),
                    c(lambda_burk$lambda %>% round(3), lambda_burk$P %>% round(4),
                      "Gram Negative", "Proteobacteria", "Beta", "Burkholderiales", "All"),
                    c(lambda_comamo$lambda %>% round(3), lambda_comamo$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Beta", "Burkholderiales", "Comamonadaceae"),
                    c(lambda_gamma$lambda %>% round(3), lambda_gamma$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Gamma", "All", "All"),
                    c(lambda_xantho$lambda %>% round(3), lambda_xantho$P %>% round(4), 
                      "Gram Negative", "Proteobacteria", "Gamma", "Xanthomonadales", "All")) %>% 
  set_colnames(c("lambda","p","Gram", "Phylum","Class","Order","Family") ) %>% 
  as.data.frame() %>% 
  dplyr::arrange(Gram, Phylum, Class, Order, Family) %>% 
  dplyr::select(Gram, Phylum, Class, Order, Family, lambda, p.value = p) %>% 
  mutate(p.value = as.numeric(p.value))
```


```{r}
lambda_tab %>% 
  group_by(Gram, Phylum) %>% 
  gt::gt() %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.05),
    fn = function(x) paste(x, "*")
  ) %>% 
    gt::text_transform(
    locations = gt::cells_body(
      columns = "p.value",
      rows = p.value < 0.01),
    fn = function(x) paste(x, "*")
  ) %>% gt::tab_style(
    style = list(
      gt::cell_text(align = "left")
      ),
    locations = gt::cells_body(columns = "p.value")
  ) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
      ),
    locations = gt::cells_body(
      columns = c("Class","Order","Family","lambda","p.value"),
      rows = p.value < 0.05 )
  ) %>% 
  write_rds("files_publication/TabS2_lambda.rds")
```


# SessionInfo

```{r}
sessionInfo()
```